### 2013-05-06 thesis.tex
Moved motivation into motivation.tex

\documentclass{article}

%% \usepackage{sdrt}
\usepackage{gb4e}
\usepackage{bussproofs}

%% Linear implication from Alessio Guglielmi
%% https://groups.google.com/forum/?fromgroups=#!topic/comp.text.tex/0B4C3F_BsVI
\def\limp {\mathbin{{-}\mkern-3.5mu{\circ}}}

\begin{document}

\author{Ji\v{r}\'{i} Mar\v{s}\'{i}k}
\title{Master Thesis}
\maketitle

\section{Preliminaries}

\input{motivation}

\input{acg}

\bibliographystyle{plain}
\bibliography{biblio}

\end{document}