### 2013-05-21 my-work-rewritten.tex
Started rewriting my-work

\section{Wide-coverage abstract categorial grammars}

Our goal is to develop large abstract categorial grammars which cover a
variety of linguistic phenomena. All of these phenomena have been
studied in detail but we would like to have a grammar which gracefully
combines all these solutions. At this scale, working out the grammar on
paper stops being sufficient to detect all the subtle interactions and
it is desirable to have a formalized representation that can be reasoned
about or tested by a computer.

To some extent, this need is already met by the \textbf{ACG development
  toolkit}\footnote{\url{http://www.loria.fr/equipes/calligramme/acg/}}. The
toolkit reads in signature and lexicon definitions and offers two modes
of operation. In the first, it checks the validity of the defined
ACG. That is to say, it checks whether the terms assigned by the
lexicons are well-typed and whether their types are consistent with the
object types dictated by the lexicon. The second mode of operation
offers an interactive experience in which the user is free to run type
inference on terms built on one of their signatures to find out whether
they are well-typed and if so, what is their principal type. The
interactive mode then lets the user apply a lexicon, or a composition of
lexicons, to a term and get the $\beta$-normal form of the object term.

We consider the above capabilities vital for doing any involved grammar
engineering in the framework of ACGs. However, there are some
limitations which make it not sufficient for our needs. The foremost of
those is that the signatures and lexicons are all explicitly realized in
(primary) memory and the toolkit expects them to be given directly in a
file. The grammars that we would like to develop will cover an
exhaustive list of wordforms and the size of our grammar definitions
would therefore grow proportionately with the size of our dictionary.

Given this setup, there are several approaches we might try, none of
them too appealing. We might write our metagrammar in a different format
and use a tool which would combine the metagrammar and the lexical
database\footnote{We will be using the term \emph{lexical database} to
  refer to what is usually called a lexicon, in order not to cause
  confusion with the lexicons of ACGs.} to produce the lexicalized ACG
in a direct representation, ready to be loaded by the toolkit. This
workflow would make development on the grammar tedious as every time we
would change the metagrammar, we would have to regenerate the direct
representation and then load it in the ACG toolkit. With extra effort,
the former cost could be alleviated by devising a system for
regenerating only parts of the grammars which will have changed since
last time. However, we would still have to pay the price of loading the
entire grammar into the toolkit before being able to interact with it
again.

Furthermore, we could pick a small representative fragment of the
dictionary and test only on that fragment during the development of our
grammar, to reduce the time it takes to both generate the direct
representation of the grammar and the time it takes to load it in the
toolkit. This would have made developing the grammar already
tenable.

However, in our approach, we have opted to spend more time on tool
development to ease subsequent work and we have developed a system which
makes defining and experimenting with lexicalized ACGs easier. The
metagrammars are defined as relational procedures which link the items
of the lexical database to elements of signatures and lexicons. The
grammars can be redefined without having to reload the lexical database
or any file of similar scale. Type inference and lexicon application
with $\beta$-normalization are provided as relations which can be used
interactively or as part of a larger program. En example of such a
larger program might be a routine to check the validity of an ACG or to
test some of its properties.

In this chapter, we will spend time explaining the decisions which went
into designing the system and the form it has now.

\subsection{}