### 2013-06-13 thesis.tex
Writing about GACGs...

\documentclass{article}

%% \usepackage{sdrt}
\usepackage{amsmath}
\usepackage{xcolor}
\usepackage{graphicx}
\usepackage{gb4e}
\usepackage{bussproofs}
\usepackage{hyperref}
\usepackage[yyyymmdd,hhmmss]{datetime}

%% Linear implication from Alessio Guglielmi
%% https://groups.google.com/forum/?fromgroups=#!topic/comp.text.tex/0B4C3F_BsVI
\def\limp {\mathbin{{-}\mkern-3.5mu{\circ}}}

\begin{document}

\author{Ji\v{r}\'{i} Mar\v{s}\'{i}k}
\title{Master Thesis}
\maketitle

\currenttime

\tableofcontents

\section{Preliminaries}

\input{motivation}

\input{acg}

\input{ig}

\input{my-work-rewritten}

\input{the-last-part}

\bibliographystyle{plain}
\bibliography{biblio}

\end{document}