### 2013-06-16 the-last-part.tex
Finished the graphical ACG chapter

\chapter{The Last Part}

In this chapter, we will attempt to incorporate accounts of several
linguistic phenomena in a single framework, highlight the modularity
issues this brings up and propose some solutions to them.

\section{Negation}
\label{sec:negation}

We will start with an account of the paired grammatical words mechanism
for negation in French. Recall the EPTDs we have seen
in~\ref{ssec:frigram}, repeated here on Figure~\ref{fig:ig-neg-rep}.

\begin{figure}
  \centering
  \includegraphics[scale=0.25]{ig-neg.pdf}
  \caption{\label{fig:ig-neg-rep} Frigram EPTDs for a negative
    determiner and the paired grammatical word $ne$.}
\end{figure}

At a basic level, $aucun$ functions like any other determiner and $ne$
as any verb modifier. However, we want to encode the constraint that
whenever a verb is modified by $ne$, the modified verb will demand that
one of its arguments contains a negative determiner. Since we need to
make a distinction between phrases that contain negative determiners and
those that do not, we will need to provide different types for both (two
terms with the same type are indistinguishable w.r.t. syntactic
composition in a type-logical grammar such as ACGs). Our type for
$aucun$ will thus end up being $N\_NEG{=}F \limp ((NP\_NEG{=}T \limp S)
\limp S)$ instead of the usual $N \limp ((NP \limp S) \limp
S)$.\footnote{Strictly following the EPTD for $aucun$ would lead us to
  give it a type $DET\_NEG{=}F$ and then give the two types
  $DET\_NEG{=}F \limp ((NP\_NEG{=}F \limp S) \limp S)$ and $DET\_NEG{=}T
  \limp ((NP\_NEG{=}T \limp S) \limp S)$ to nouns. In this
  demonstration, we prefer to align our examples with the treatment
  which is customary to ACG research.} The type for $le$ will need to be
able handle cases where its argument noun either already contains a
negative determiner or not.

\begin{align*}
N_{aucun} &: N\_NEG{=}F \limp ((NP\_NEG{=}T \limp S) \limp S) \\
N_{le_1} &: N\_NEG{=}F \limp ((NP\_NEG{=}F \limp S) \limp S) \\
N_{le_2} &: N\_NEG{=}T \limp ((NP\_NEG{=}T \limp S) \limp S) \\
\end{align*}

The type of $ne$ will be more verbose, as it will consume a verb and
transform its valency so that it demands that exactly one of its $NP$
arguments contains a negative determiner.

\begin{align*}
N_{ne_{tv_1}} &: (NP\_NEG{=}F \limp NP\_NEG{=}F \limp S) \limp (NP\_NEG{=}T \limp NP\_NEG{=}F \limp S) \\
N_{ne_{tv_2}} &: (NP\_NEG{=}F \limp NP\_NEG{=}F \limp S) \limp (NP\_NEG{=}F \limp NP\_NEG{=}T \limp S)
\end{align*}

Furthermore, we could add the case when both the subject and the object
contain negative determiners, such as in~(\ref{ex:aucun2}). This case is
currently not covered by Frigram (though it could be), but is considered
grammatical by French speakers. Finally, in a complete grammar, $ne$
would have to provide types capable of transforming all the other
syntactic valencies.

\begin{exe}
  \ex \label{ex:aucun2} Aucune fourmi n'aime aucun tatou.
\end{exe}

\begin{align*}
N_{ne_{tv_3}} &: (NP\_NEG{=}F \limp NP\_NEG{=}F \limp S) \limp (NP\_NEG{=}T \limp NP\_NEG{=}T \limp S)
\end{align*}

\begin{figure}
  \centering
  \includegraphics[scale=0.25]{parse-aucun.pdf}
  \caption{\label{fig:parse-aucun} One of the parse trees assigned to
    sentence (\ref{ex:aucun-obj}) by Frigram/Leopar.}
\end{figure}

However, this is not the only change we have to effect on our grammar in
order to properly handle the paired grammatical words for
negation. Consider the example sentence (\ref{ex:aucun-obj}) and the
parse tree assigned to it by Frigram on Figure~\ref{fig:parse-aucun}. It
is not enough that the embedded noun phrase $aucun\ tatou$ has a type
that tells us it contains a negative determiner. We need this bit of
internal information also at the level of the enclosing noun phrase
$l'odeur\ d'aucun\ tatou$ in which the former noun phrase is embedded
since this is a property of the enclosing noun phrase that is crucial to
its syntactic combinatorics. This presupposes the existence of a
mechanism for propagating this kind of information up the parse tree. In
the case of our ACG, this means that all the functions which modify
phrases must be aware whether or not their argument contains a negative
determiner and to convey this information in its result type as
well. Moreover, since the negative determiner can also come from a
complement of a prepositional phrase, the types of prepositions will
need to take into account the presence of negative determiners in both
their complements and the phrases they modify.

\begin{align*}
N_{que_1} &: (NP\_NEG{=}F \limp S) \limp N\_NEG{=}F \limp N\_NEG{=}F \\
N_{que_2} &: (NP\_NEG{=}F \limp S) \limp N\_NEG{=}T \limp N\_NEG{=}T \\
N_{de_1} &: NP\_NEG{=}F \limp N\_NEG{=}F \limp N\_NEG{=}F \\
N_{de_2} &: NP\_NEG{=}F \limp N\_NEG{=}T \limp N\_NEG{=}T \\
N_{de_3} &: NP\_NEG{=}T \limp N\_NEG{=}F \limp N\_NEG{=}T
\end{align*}

If we suppose the presence of the appropriate lexical items typed as in
the examples of~\ref{sec:acg}, we have a type system in which we can
type the syntactic terms corresponding to the sentences
(\ref{ex:good-neg-double}), (\ref{ex:good-neg-embed}) and
(\ref{ex:good-neg-rel}) while making it impossible to type the
structures corresponding to sentences (\ref{ex:bad-neg-noneg}) and
({\ref{ex:bad-neg-rel}}).

In sentence (\ref{ex:bad-neg-noneg}), we have no constant
$N_{ne_{tv_?}}$ for $ne$ which would yield a verb capable of consuming
two noun phrases that contain no negative determiners.

In sentence (\ref{ex:bad-neg-rel}), there are multiple problems with
typing. First, $N_{chasse}$ requires its first argument to be of type
$NP\_NEG{=}F$, a noun phrase with no negative determiner. This means
that $\lambda^{\circ} x.\ N_{chasse}\ x\ y$ has type $NP\_NEG{=}F \limp
S$ which makes it an invalid argument for the function
$N_{aucun}\ N_{loup}$, which has type $(NP\_NEG{=}T \limp S) \limp
S$. Even if we were able to assign it a type compatible with
$N_{que_1}$, the final relative clause would map $N_{tatou}$ to another
term of type $N\_NEG{=}F$, which has no ``free'' negative
determiners. $N_{le_1}$ will map this to a term typed $(NP\_NEG{=}F
\limp S) \limp S$. Finally, this term is not capable of taking
$(\lambda^{\circ} x.\ N_{ne_{iv}}\ N_{court}\ x)$ as its argument, since
its type is $NP\_NEG{=}T \limp S$.

\begin{exe}
  \ex \label{ex:good-neg-double} Aucune fourmi n'aime aucun tatou. \\
      $(N_{aucune}\ N_{fourmi})\ (\lambda^{\circ} x.\ (N_{aucun}\ N_{tatou})\ (\lambda^{\circ} y.\ N_{ne_{tv_3}}\ N_{aime}\ x\ y))$
  \ex \label{ex:good-neg-embed} Jean n'aime l'odeur d'aucun tatou. \\
      $(N_{le_2}\ (N_{de_3}\ (N_{aucun}\ N_{tatou})\ N_{odeur}))\ (\lambda^{\circ} y.\ N_{ne_{tv_2}}\ N_{aime}\ N_{Jean}\ y)$
  \ex \label{ex:good-neg-rel} Le tatou qu'aucun loup ne chasse court. \\
      $(N_{le_1}\ (N_{que_1}\ (\lambda^{\circ} y.\ (N_{aucun}\ N_{loup})\ (\lambda^{\circ} x.\ N_{ne_{tv_1}}\ N_{chasse}\ x\ y))\ N_{tatou}))\ (\lambda^{\circ} x.\ N_{court}\ x)$
  \ex * \label{ex:bad-neg-noneg} Jean n'aime le tatou. \\
      * $(N_{le_1}\ N_{tatou})\ (\lambda^{\circ} y.\ N_{ne_{tv_?}}\ N_{aime}\ N_{Jean}\ y)$
  \ex * \label{ex:bad-neg-rel} Le tatou qu'aucun loup chasse ne court. \\
      * $(N_{le_1}\ (N_{que}\ (\lambda^{\circ} y.\ (N_{aucun}\ N_{loup})\ (\lambda^{\circ} x.\ N_{chasse}\ x\ y))\ N_{tatou}))\ (\lambda^{\circ} x.\ N_{ne_{iv}}\ N_{court}\ x)$
\end{exe}

There are two things that we would like to highlight about this
grammatical treatment. First off, we had to refine our types to make
them convey more than one piece of information in a single type
(e.g. something is a noun phrase and at the same time it is something
that contains a negative determiner). Second, a property that we wish to
express in the type of a term may be due to one of its subconstituents
and it is therefore necessary to encode the propagation of this property
up the parse tree in the types of all the intervening operators. In our
case, it meant that handling negative determiners and the paired
grammatical word for negation required us to not only define and type
constants for these two categories, but to also change the types of noun
modifiers ($que\ ...$) and prepositions ($de$).

\section{Extraction}
\label{sec:extraction}

We will now turn to another phenomenon, extraction in relative clauses,
focus on one of its properties, the fact that subjects can be extracted
using $qui$ only from the relative clause itself whereas objects can be
extracted using $que$ from embedded clauses contained therein, and
present an implementation of this constraint by way of types in an
ACG. The solution that we will show was published in
\cite{pogodalla2012controlling}. Our presentation is slightly adapted to
conform to the patterns seen in previous examples and it does not use
dependent types as they are not essential to this specific constraint.

The key idea in the treatment presented below is to give a special type
to extracted constituents. We will thus have two types for noun phrases:
$NP\_VAR{=}T$, which stands for ``empty'' noun phrases introduced by
extraction, and $NP\_VAR{=}F$, which stands for the other noun phrases,
i.e.  proper names and nouns with determiners. The grammar gives no
constant capable of constructing a value of type $NP\_VAR{=}T$. The only
way to obtain a term with this type is to abstract over a variable
having the type. Then, the types of relative pronouns end up being
$(NP\_VAR{=}T \limp S) \limp N \limp N$, ensuring that the incomplete
clause given as the first argument is a $\lambda$-abstraction whose
argument has type $NP\_VAR{=}T$.

Once we have this kind of information in the type system, we can start
distinguishing clauses containing rooted extraction, where a subject or
object is extracted directly from the relative clause, embedded
extraction, where an object is extracted from a clause embedded inside
the relative clause, or no extraction. Types of functions which produce
values of type $S$ will be sensitive to the presence of extracted
variables in their noun phrase and clausal arguments.

\begin{align*}
E_{dort_1} &: NP\_VAR{=}F \limp S\_EXT{=}NO \\
E_{dort_2} &: NP\_VAR{=}T \limp S\_EXT{=}ROOT \\
E_{aime_1} &: NP\_VAR{=}F \limp NP\_VAR{=}F \limp S\_EXT{=}NO \\
E_{aime_2} &: NP\_VAR{=}T \limp NP\_VAR{=}F \limp S\_EXT{=}ROOT \\
E_{aime_3} &: NP\_VAR{=}F \limp NP\_VAR{=}T \limp S\_EXT{=}ROOT \\
E_{dit\ que_1} &: NP\_VAR{=}F \limp S\_EXT{=}NO \limp S\_EXT{=}NO \\
E_{dit\ que_2} &: NP\_VAR{=}T \limp S\_EXT{=}NO \limp S\_EXT{=}ROOT \\
E_{dit\ que_3} &: NP\_VAR{=}F \limp S\_EXT{=}ROOT \limp S\_EXT{=}EMB \\
E_{dit\ que_4} &: NP\_VAR{=}F \limp S\_EXT{=}EMB \limp S\_EXT{=}EMB
\end{align*}

Now all of the infrastructure is in place for us to express the type of
the relative pronouns $qui$ and $que$.

\begin{align*}
E_{qui} &: (NP\_VAR{=}T \limp S\_EXT{=}ROOT) \limp N \limp N \\
E_{que_1} &: (NP\_VAR{=}T \limp S\_EXT{=}ROOT) \limp N \limp N \\
E_{que_2} &: (NP\_VAR{=}T \limp S\_EXT{=}EMB) \limp N \limp N
\end{align*}

The above types ensure that the abstracted variables corresponding to
traces of extracted constituents will be given the type $NP\_VAR{=}T$,
that $qui$ can only be used to extract constituents from root positions
while $que$ allows for extraction from both root and embedded
positions.\footnote{NB: The constraint that $qui$ can only extract
  subjects and that $que$ can only extract objects is not enforced
  here. Likewise, this fragment does not handle multiple extraction.}

We also note that the splitting of the clause type $S$ into three finer
types $S\_EXT={NO}$, $S\_EXT{=}ROOT$ and $S\_EXT{=}EMB$ means that types
that work with $S$ regardless of its extraction status will have to
provide alternatives for all the possible variants.

\begin{align*}
E_{le_1} &: N \limp ((NP\_VAR{=}F \limp S\_EXT{=}NO) \limp S\_EXT{=}NO) \\
E_{le_2} &: N \limp ((NP\_VAR{=}F \limp S\_EXT{=}ROOT) \limp S\_EXT{=}ROOT) \\
E_{le_3} &: N \limp ((NP\_VAR{=}F \limp S\_EXT{=}EMB) \limp S\_EXT{=}EMB)
\end{align*}

While we have introduced the notion of an $NP\_VAR{=}T$ for describing
traces, we can cover another element of our running fragment, the
preposition $de$, and enforce the constraint that $qui$/$que$ cannot
extract prepositional complements by providing only a single type for
$de$ that accepts only unmoved noun phrases as complements.

\begin{align*}
E_{de} &: NP\_VAR{=}F \limp N \limp N
\end{align*}

Given this type system, we can assign the type $S\_EXT{=}NO$ to the term
expressing the syntactic structure of (\ref{ex:good-ext}) but we cannot
a find a typable term for sentence (\ref{ex:bad-ext}), where $qui$ is
used to extract a subject from a clause embedded in a relative clause.
The problem with sentence (\ref{ex:bad-ext}) is that the type of the
relative clause is $NP\_VAR{=}T \limp S\_EXT{=}EMB$ which is not a valid
argument for any constant representing the relative pronoun
$qui$. Typing the variable $t$ with $NP\_VAR{=}F$ instead of
$NP\_VAR{=}T$ and using $E_{aime_1}$ instead of $E_{aime_2}$ would not
help, since the resulting relative clause would then have type
$NP\_VAR{=}F \limp S\_EXT{=}NO$, which would not be admitted by any
relative pronoun.

\begin{exe}
  \ex * \label{ex:bad-ext} Le tatou qui Jean dit que \_ aime Marie dort. \\
      * $(E_{le_1}\ (E_{qui_?}\ (\lambda^{\circ} t.\ E_{dit\ que_3}\ E_{Jean}\ (E_{aime_2}\ t\ E_{Marie}))\ E_{tatou}))\ (\lambda^{\circ} x.\ E_{dort_1}\ x)$
  \ex \label{ex:good-ext} Le tatou que Jean dit que Marie aime \_ dort. \\
      $(E_{le_1}\ (E_{que_2}\ (\lambda^{\circ} t.\ E_{dit\ que_3}\ E_{Jean}\ (E_{aime_3}\ E_{Marie}\ t))\ E_{tatou}))\ (\lambda^{\circ} x.\ E_{dort_1}\ x)$
\end{exe}

As was the case with our treatment of negation, we refined our familiar
types by tacking on new kinds of information and defined rules for how
these pieces of information percolate up the parse tree all the way to
the topmost level still pertinent for ensuring grammaticality.

\section{Agreement}
\label{sec:agreement}

Finally, we will look at another example of implementing a linguistic
constraint in ACGs by covering the phenomenon of number
agreement.

Number agreement motivates a straightforward refinement of our types,
where all noun and noun phrase types carry number information.

\begin{align*}
A_{Marie} &: NP\_NUM{=}SG \\
A_{tatou} &: N\_NUM{=}SG \\
A_{tatous} &: N\_NUM{=}PL \\
A_{le} &: N\_NUM{=}SG \limp ((NP\_NUM{=}SG \limp S) \limp S) \\
A_{les} &: N\_NUM{=}PL \limp ((NP\_NUM{=}PL \limp S) \limp S) \\
A_{dort} &: NP\_NUM{=}SG \limp S \\
A_{dorment} &: NP\_NUM{=}PL \limp S \\
A_{aime_1} &: NP\_NUM{=}SG \limp NP\_NUM{=}SG \limp S \\
A_{aime_2} &: NP\_NUM{=}SG \limp NP\_NUM{=}PL \limp S \\
A_{aiment_1} &: NP\_NUM{=}PL \limp NP\_NUM{=}SG \limp S \\
A_{aiment_2} &: NP\_NUM{=}PL \limp NP\_NUM{=}PL \limp S \\
A_{qui_1} &: (NP\_NUM{=}SG \limp S) \limp N\_NUM{=}SG \limp N\_NUM{=}SG \\
A_{qui_2} &: (NP\_NUM{=}PL \limp S) \limp N\_NUM{=}PL \limp N\_NUM{=}PL \\
A_{de_1} &: NP\_NUM{=}SG \limp N\_NUM{=}SG \limp N\_NUM{=}SG \\
A_{de_2} &: NP\_NUM{=}PL \limp N\_NUM{=}SG \limp N\_NUM{=}SG \\
A_{de_3} &: NP\_NUM{=}SG \limp N\_NUM{=}PL \limp N\_NUM{=}PL \\
A_{de_4} &: NP\_NUM{=}PL \limp N\_Pl \limp N\_NUM{=}PL
\end{align*}

Given the above signature, we can assign the type $S$ to the syntactic
term of sentence (\ref{ex:good-agr}) but not to the one of sentence
(\ref{ex:bad-agr}). In the latter sentence, the type of the variable $t$
has to be $NP\_NUM{=}SG$ since it is used as an argument to $A_{dort}$. This
means that the relative clause $dort$ has type $NP\_NUM{=}SG \limp
S$. However, there is no type for $qui$ which would allow us to modify
the plural noun $A_{tatous}$ of type $N\_NUM{=}PL$ by a relative clause with a
missing singular noun phrase, which has the type $NP\_NUM{=}SG \limp S$.

\begin{exe}
  \ex \label{ex:good-agr} Les tatous de Marie dorment. \\
      $(A_{les}\ (A_{de}\ A_{Marie}\ A_{tatous}))\ (\lambda^{\circ} x.\ A_{dorment}\ x)$
  \ex * \label{ex:bad-agr} Marie aime les tatous qui dort. \\
      * $(A_{les}\ (A_{qui_?}\ (\lambda^{\circ} t.\ A_{dort}\ t)\ A_{tatous}))\ (\lambda^{\circ} y.\ A_{aime_2}\ A_{Marie}\ y)$
\end{exe}

Once again, we have expressed a linguistic constraint by refining our
types, augmenting them with a new bit of information, we constrained the
arguments of functions using these new types and we have described how
this new type information propagates from the given arguments to the
produced values.

\section{Putting It All Together}
\label{sec:together}

In this section, we will see what it takes to enforce all of the
constraints introduced in the preceding sections (negation, extraction
and agreement) at the same time, which will give us some idea on how to
go about building a grammar which handles a wide variety of such
phenomena.

In a grammar that handles negation, extraction and agreement, our types
will need to carry all of the refinements we introduced before. To
recapitulate, $NP$ will be parameterized by whether or not it contains a
negative determiner ($\_\textcolor{red}{NEG{=}T}$ and
$\_\textcolor{red}{NEG{=}F}$ respectively), whether or not it is a trace
from an extraction ($\_\textcolor{green}{VAR{=}T}$ and
$\_\textcolor{green}{VAR{=}F}$) and its number
($\_\textcolor{blue}{NUM{=}SG}$ or $\_\textcolor{blue}{NUM{=}PL}$). For
$N$, we have the same refinements except for $\textcolor{green}{VAR}$
since extraction only moves $NP$s. Clauses, of type $S$, will be
discriminated based on the level of extraction within them
($\_\textcolor{green}{EXT{=}NO}$, $\_\textcolor{green}{EXT{=}ROOT}$ and
$\_\textcolor{green}{EXT{=}EMB}$).

The individual types will have to respect all of the constraints at the
same time, each one being a complete specification of a possible
situation (complete w.r.t. the features listed above). To see and
appreciate what this means in practice, we give the types for the
preposition $de$:

\begin{align*}
C_{de_1} &: (NP\_\textcolor{red}{NEG{=}F}\_\textcolor{green}{VAR{=}F}\_\textcolor{blue}{NUM{=}SG}) \limp (N\_\textcolor{red}{NEG{=}F}\_\textcolor{blue}{NUM{=}SG}) \limp
(N\_\textcolor{red}{NEG{=}F}\_\textcolor{blue}{NUM{=}SG}) \\
C_{de_2} &: (NP\_\textcolor{red}{NEG{=}F}\_\textcolor{green}{VAR{=}F}\_\textcolor{blue}{NUM{=}SG}) \limp (N\_\textcolor{red}{NEG{=}F}\_\textcolor{blue}{NUM{=}PL}) \limp
(N\_\textcolor{red}{NEG{=}F}\_\textcolor{blue}{NUM{=}PL}) \\
C_{de_3} &: (NP\_\textcolor{red}{NEG{=}F}\_\textcolor{green}{VAR{=}F}\_\textcolor{blue}{NUM{=}SG}) \limp (N\_\textcolor{red}{NEG{=}T}\_\textcolor{blue}{NUM{=}SG}) \limp
(N\_\textcolor{red}{NEG{=}T}\_\textcolor{blue}{NUM{=}SG}) \\
C_{de_4} &: (NP\_\textcolor{red}{NEG{=}F}\_\textcolor{green}{VAR{=}F}\_\textcolor{blue}{NUM{=}SG}) \limp (N\_\textcolor{red}{NEG{=}T}\_\textcolor{blue}{NUM{=}PL}) \limp
(N\_\textcolor{red}{NEG{=}T}\_\textcolor{blue}{NUM{=}PL}) \\
C_{de_5} &: (NP\_\textcolor{red}{NEG{=}T}\_\textcolor{green}{VAR{=}F}\_\textcolor{blue}{NUM{=}SG}) \limp (N\_\textcolor{red}{NEG{=}F}\_\textcolor{blue}{NUM{=}SG}) \limp
(N\_\textcolor{red}{NEG{=}T}\_\textcolor{blue}{NUM{=}SG}) \\
C_{de_6} &: (NP\_\textcolor{red}{NEG{=}T}\_\textcolor{green}{VAR{=}F}\_\textcolor{blue}{NUM{=}SG}) \limp (N\_\textcolor{red}{NEG{=}F}\_\textcolor{blue}{NUM{=}PL}) \limp
(N\_\textcolor{red}{NEG{=}T}\_\textcolor{blue}{NUM{=}PL}) \\
C_{de_7} &: (NP\_\textcolor{red}{NEG{=}F}\_\textcolor{green}{VAR{=}F}\_\textcolor{blue}{NUM{=}PL}) \limp (N\_\textcolor{red}{NEG{=}F}\_\textcolor{blue}{NUM{=}SG}) \limp
(N\_\textcolor{red}{NEG{=}F}\_\textcolor{blue}{NUM{=}SG}) \\
C_{de_8} &: (NP\_\textcolor{red}{NEG{=}F}\_\textcolor{green}{VAR{=}F}\_\textcolor{blue}{NUM{=}PL}) \limp (N\_\textcolor{red}{NEG{=}F}\_\textcolor{blue}{NUM{=}PL}) \limp
(N\_\textcolor{red}{NEG{=}F}\_\textcolor{blue}{NUM{=}PL}) \\
C_{de_9} &: (NP\_\textcolor{red}{NEG{=}F}\_\textcolor{green}{VAR{=}F}\_\textcolor{blue}{NUM{=}PL}) \limp (N\_\textcolor{red}{NEG{=}T}\_\textcolor{blue}{NUM{=}SG}) \limp
(N\_\textcolor{red}{NEG{=}T}\_\textcolor{blue}{NUM{=}SG}) \\
C_{de_{10}} &: (NP\_\textcolor{red}{NEG{=}F}\_\textcolor{green}{VAR{=}F}\_\textcolor{blue}{NUM{=}PL}) \limp (N\_\textcolor{red}{NEG{=}T}\_\textcolor{blue}{NUM{=}PL}) \limp
(N\_\textcolor{red}{NEG{=}T}\_\textcolor{blue}{NUM{=}PL}) \\
C_{de_{11}} &: (NP\_\textcolor{red}{NEG{=}T}\_\textcolor{green}{VAR{=}F}\_\textcolor{blue}{NUM{=}PL}) \limp (N\_\textcolor{red}{NEG{=}F}\_\textcolor{blue}{NUM{=}SG}) \limp
(N\_\textcolor{red}{NEG{=}T}\_\textcolor{blue}{NUM{=}SG}) \\
C_{de_{12}} &: (NP\_\textcolor{red}{NEG{=}T}\_\textcolor{green}{VAR{=}F}\_\textcolor{blue}{NUM{=}PL}) \limp (N\_\textcolor{red}{NEG{=}F}\_\textcolor{blue}{NUM{=}PL}) \limp
(N\_\textcolor{red}{NEG{=}T}\_\textcolor{blue}{NUM{=}PL})
\end{align*}

This highlights two important problems with the current approach. One is
the untenable complexity. Even though the three mechanisms are
completely independent of each other, here we are forced to refer to all
of them in every type. It is no longer easy to glance at the types
assigned to $de$ and see how it behaves with respect to, for example,
the mechanism of negation (compare with the type assignments we gave to
$de$ in previous subsections). Another cause of unreadability of this
signature is due to the sheer number of types on display, which leads us
to our second issue.

The signature demonstrates an exponential growth in the number of typed
constants included. In cases where the behaviors of a wordform across
the different mechanisms are independent, the complete type signature
will need to provide a type for every combination of situations in all
of the mechanisms. Assuming that every mechanism will assign on average
more than one type to every wordform, the number of the wordform's types
in the final signature will be exponential w.r.t. the number of
different mechanisms handled.

We can try making the enumeration of these types easier for the grammar
author by introducing metavariables for the values of the features. This
means that we could write the following 3 templates instead of the 12
types above:

$$
\forall \textcolor{blue}{x}, \textcolor{blue}{y} \in \{\textcolor{blue}{SG}, \textcolor{blue}{PL}\}
$$
\begin{align*}
C_{de_{1(\textcolor{blue}{x}, \textcolor{blue}{y})}} &: (NP\_\textcolor{red}{NEG{=}F}\_\textcolor{green}{VAR{=}F}\_\textcolor{blue}{NUM{=}x}) \limp (N\_\textcolor{red}{NEG{=}F}\_\textcolor{blue}{NUM{=}y}) \limp (N\_\textcolor{red}{NEG{=}F}\_\textcolor{blue}{NUM{=}y}) \\
C_{de_{2(\textcolor{blue}{x}, \textcolor{blue}{y})}} &: (NP\_\textcolor{red}{NEG{=}F}\_\textcolor{green}{VAR{=}F}\_\textcolor{blue}{NUM{=}x}) \limp (N\_\textcolor{red}{NEG{=}T}\_\textcolor{blue}{NUM{=}y}) \limp (N\_\textcolor{red}{NEG{=}T}\_\textcolor{blue}{NUM{=}y}) \\
C_{de_{3(\textcolor{blue}{x}, \textcolor{blue}{y})}} &: (NP\_\textcolor{red}{NEG{=}T}\_\textcolor{green}{VAR{=}F}\_\textcolor{blue}{NUM{=}x}) \limp (N\_\textcolor{red}{NEG{=}F}\_\textcolor{blue}{NUM{=}y}) \limp (N\_\textcolor{red}{NEG{=}T}\_\textcolor{blue}{NUM{=}y})
\end{align*}

This is getting more concise and easier to comprehend. We can take this
approach further by not only quantifying over variables but also
computing some of the values using functions:

$$
\forall \textcolor{blue}{x}, \textcolor{blue}{y} \in\{\textcolor{blue}{SG}, \textcolor{blue}{PL}\}, \forall \textcolor{red}{p}, \textcolor{red}{q} \in \{\textcolor{red}{F}, \textcolor{red}{T}\}, \textcolor{red}{p \land q} \neq \textcolor{red}{T}
$$
\begin{align*}
C_{de(\textcolor{blue}{x}, \textcolor{blue}{y}, \textcolor{red}{p}, \textcolor{red}{q})} &: (NP\_\textcolor{red}{NEG{=}p}\_\textcolor{green}{VAR{=}F}\_\textcolor{blue}{NUM{=}x}) \limp (N\_\textcolor{red}{NEG{=}q}\_\textcolor{blue}{NUM{=}y}) \limp (N\_\textcolor{red}{NEG{=}(p \lor q)}\_\textcolor{blue}{NUM{=}y})
\end{align*}

The above is vastly more concise and manages to convey the intent more
clearly (the intent being that it cannot be the case that there are
negative determiners in both the complement noun phrase and the modified
noun and that the value of the result's \textcolor{red}{NEG} feature is
always the disjunction of the \textcolor{red}{NEG} features of the two
arguments).

As of right now, we have introduced this kind of rule only as
meta-notation. This means that the real underlying grammar that will end
up being used by some algorithm still suffers from the same exponential
cardinality of the signature. Fortunately, there exists a well-studied
construction in type theory that lets us assign to terms generic types
such as the one defined in the template above. This construction is the
\emph{dependent product} and its utility in ACGs has already been
noted.\footnote{See \cite{de2007two} for its introduction into the
  domain of ACGs, \cite{de2007type} for a small example grammar, which
  handles the interaction between number and gender agreement and
  coordination, and finally see \cite{pogodalla2012controlling} for
  using dependent types to handle several constraints regarding
  extraction (including the one we covered in~\ref{sec:extraction}).}

While extending the type system by including dependent types is
definitely useful and it helps to solve the problem of having our
signatures grow exponentially in size, it does not solve our problem
with the complexity of the new grammar. Granted, the new meta-type
written in the dependent style is more readable and easier to reason
about than the enumeration of 12 basic types we presented at the
beginning of this section, however, the result is still complex in
that it forces us to describe the behavior of three independent
mechanisms in one place. This one place happens to be a wordform which
has seemingly little connection to these mechanisms, the preposition
$de$. We can therefore imagine that it will participate in many other
phenomena and that the type we assign to it will grow beyond our
abilities to reason about it effectively.

Furthermore, the current notation does not even make it obvious that
these are independent mechanisms. One has to examine the type to make
sure that there is no interaction. In the rest of this chapter, we will
devote ourselves to investigating ways of disentangling these.

\section{Intersections of Types}

Our goal is to split the types into smaller parts that only talk about
features belonging to one of the three mechanisms covered above. This
means that we would like to take some phrase about which we know a lot
of disparate pieces of information (presence of negative determiner,
trace status, number\ldots), having type
e.g. $NP\_\textcolor{red}{NEG{=}T}\_\textcolor{green}{VAR{=}F}\_\textcolor{blue}{NUM{=}PL}$,
and use it as an argument for some function which only cares about one
channel of information and would have a type like
$NP\_\textcolor{red}{NEG{=}T} \limp N\_\textcolor{red}{NEG{=}F} \limp
N\_\textcolor{red}{NEG{=}T}$. To make a type such as
$NP\_\textcolor{red}{NEG{=}T}\_\textcolor{green}{VAR{=}F}\_\textcolor{blue}{NUM{=}PL}$
be a legal argument to a function which expects an argument of type
$NP\_\textcolor{red}{NEG{=}T}$, we will need to introduce some notion of
\emph{subtyping}.

The types we have been showing up to now propose a very straightforward
notion of subtyping. All of the types can be seen as conjunctions of
properties where $\_$ (underscore) is the conjunction operator. A type
$\tau$ can then be seen as a subtype of type $\sigma$ whenever the
conjuncts of $\tau$ form a superset of the conjuncts of $\sigma$. This
is not unlike the structural typing of records in OCaml (CITE SOME OCAML
SPEC HERE), where a record type is a subtype of another record type if
the former provides, among others, all the fields provided by the latter
and gives them the same type. In our case, one type is a subtype of
another if the former gives, among others, all the properties given by
the latter. This notion of subtyping can be expressed by adding the rule
of inference in Figure~\ref{fig:type-superset} to our type
system.\footnote{In the rule \ldots (talk about $\land$ instead of $\_$
  (underscore) here...)}

\begin{figure}
  \begin{prooftree}
    \AxiomC{$\Gamma_l; \Gamma_i \vdash_\Sigma M : A_1 \land \ldots A_{i-1} \land A_i \land A_{i+1} \ldots \land A_n$}
    \RightLabel{(subtype)}
    \UnaryInfC{$\Gamma_l; \Gamma_i \vdash_\Sigma M : A_1 \land \ldots A_{i-1} \land A_{i+1} \ldots \land A_n$}
  \end{prooftree}
  \caption{\label{fig:type-superset} A rule of inference giving us
    subtyping for conjunctions of atomic types. For readability, we use
    $\land$ instead of $\_$ (underscore) to designate the conjunction of
    properties.}
\end{figure}

(Talk about whether or not this is a conservative extension...)

With the new subtyping rule, we can

$$
S_{tatou} : N \land NEG{=}F \land NUM{=}SG
$$
$$
S_{le_{NEG_1}} : (N \land NEG{=}F) \limp (((NP \land NEG{=}F) \limp S) \limp S)
$$
$$
S_{le_{NEG_2}} : (N \land NEG{=}T) \limp (((NP \land NEG{=}T) \limp S) \limp S)
$$
$$
S_{le_{NUM}} : (N \land NUM{=}SG) \limp (((NP \land NUM{=}SG) \limp S) \limp S)
$$
$$
S_{le_{VAR}} : N \limp (((NP \land VAR{=}F) \limp S) \limp S)
$$

\ldots

\section{Intersections of Languages}
\label{sec:sects-of-langs}

We will turn to another solution to the problem of complex grammars that
try to implement multiple independent constraints. This solution will be
based on a generalization of ACGs, but in order to introduce this
generalization, we will first need to present the manner in which ACGs
and their compositions are formulated in ACG literature.

\subsection{The Usual Formalization of ACGs}
\label{ssec:usual-acgs}

Let us consider a system with three signatures $\Sigma_1$, $\Sigma_2$
and $\Sigma_3$, their respective distinguished types $S_1$, $S_2$ and
$S_3$ and two lexicons $\mathcal{L}_{12} : \Sigma_1 \to \Sigma_2$ and
$\mathcal{L}_{23} : \Sigma_2 \to \Sigma_3$ such that
$\mathcal{L}_{12}(S_1) = S_2$ and $\mathcal{L}_{23}(S_2) = S_3$. See
Figure~\ref{fig:acg-serial-comp} for a diagrammatic visualization of
this structure.

\begin{figure}[t]
  \centering
  \begin{subfigure}[b]{0.4\textwidth}
    \centering
    \includegraphics[height=0.2\textheight]{diagrams/acg-serial-comp.pdf}
    \caption{\label{fig:acg-serial-comp} Serial composition of ACGs.}
  \end{subfigure}
  \qquad
  \begin{subfigure}[b]{0.4\textwidth}
    \centering
    \includegraphics[height=0.2\textheight]{diagrams/acg-parallel-comp.pdf}
    \caption{\label{fig:acg-parallel-comp} Parallel composition of ACGs.}
  \end{subfigure}
  \caption{\label{fig:acg-comp-modes} Two modes of composition for
    ACGs.}
\end{figure}

We can now define languages $A_1$, $A_2$ and $A_3$ where $A_i$ is the
set of terms from $\Lambda(\Sigma_i)$ that have the type $S_i$. Such
languages, which are defined only in terms of a signature and some
distinguished type, are called \emph{abstract languages} in the
terminology of ACGs. They consist of terms whose structure is
constrained only by the types assigned by the signature.

Further than that, we can look at the languages $\mathcal{L}_{12}(A_1)$
and $\mathcal{L}_{23}(A_2)$, images of the abstract languages given by
the lexicons. Since the lexicons are homomorphic mappings w.r.t. the
types, all elements of $\mathcal{L}_{12}(A_1)$ will have the type
$\mathcal{L}_{12}(S_1) = S_2$ and will therefore also belong to $A_2$
(similarly, we have that $\mathcal{L}_{23}(A_2) \subseteq A_3$). These
new languages, which are formed by mapping an ACG abstract language
using a lexicon, are called \emph{object languages}. Their terms are
constrained not only by the types of the signature they are built upon
but also by the type system of the abstract signature.

Finally, we can also take the composition of lexicons $\mathcal{L}_{13}
= \mathcal{L}_{23} \circ \mathcal{L}_{12}$, which itself is also a
lexicon, and map it over $A_1$. This gives us a language
$\mathcal{L}_{13}(A_1) = \mathcal{L}_{23}(\mathcal{L}_{12}(A_1))$ which
is at the same time the image of the object language
$\mathcal{L}_{12}(A_1)$ given by $\mathcal{L}_{23}$ and also the image
of the abstract language $A_1$ given by $\mathcal{L}_{13}$ (thus it is
itself also an object language). This points to an interesting property
of ACG languages: they are closed under transformation by a lexicon.

In the ACG literature, ACGs are defined as quadruples $\mathcal{G} =
\mathopen{<}\Sigma_A, \Sigma_O, \mathcal{L}, S\mathclose{>}$ where
$\Sigma_A$ is the abstract signature, $\Sigma_O$ is the object
signature, $\mathcal{L} : \Sigma_A \to \Sigma_O$ is a lexicon linking
the two and $S$ is a distinguished type of the abstract signature
$\Sigma_A$. From this grammar, we can define two languages. The abstract
language $\mathcal{A}(\mathcal{G}) = \{t \in \Lambda(\Sigma_A) \mid
\ \vdash_{\Sigma_A} t : S\}$ and the object language
$\mathcal{O}(\mathcal{G}) = \mathcal{L}(\mathcal{A}(\mathcal{G})) = \{t
\in \Lambda(\Sigma_O) \mid \exists u \in \mathcal{A}(\mathcal{G}),
\mathcal{L}(u) = t\}$. The composition of two ACGs $\mathcal{G}_1 =
\mathopen{<}\Sigma_1, \Sigma_2, \mathcal{L}_{12}, S_1\mathclose{>}$ and
$\mathcal{G}_2 = \mathopen{<}\Sigma_2, \Sigma_3, \mathcal{L}_{23},
S_2\mathclose{>}$ is defined as $\mathcal{G}_2 \circ \mathcal{G}_1 =
\mathopen{<}\Sigma_1, \Sigma_3, \mathcal{L}_{23} \circ \mathcal{L}_{12},
S_1\mathclose{>}$.

The definition of ACGs given above is quite practical in that it is
complete, i.e. it gives a precise account of what set of languages are
generated by ACGs (abstract and object languages for which there exist
the quadruples described above). This contrasts with the presentation of
ACGs that we gave in~\ref{sec:acg}, where we defined ACGs as collections
of signatures associated with distinguished types and connected with
lexicons. Our definition only introduced the requisite machinery, but
delayed the delimitation of the way languages can be defined. Leaving
this question open will let us answer it now by providing an extension
of ACGs which will help us overcome the modularity issues that have been
the theme of this chapter.\footnote{Incidentally, this was not the
  reason for the way we presented ACGs in~\ref{sec:acg}. We chose this
  style as it grouped together the information in a more natural way. We
  approach signatures as descriptions of domains (strings, syntax,
  semantics), each one being associated with a distinguished type
  telling us which objects the domain has set out to describe (strings,
  sentences, truth values). Having distinguished types directly attached
  to signatures allows us to omit them from the ACG quadruples,
  $\mathopen{<}\Sigma_A, \Sigma_O, \mathcal{L},
  \cancel{S}\mathclose{>}$. However, this only leaves us with triples
  containing the two signatures and the lexicon connecting them, and
  since we can infer the signatures themselves from the domain and the
  range of the lexicon $\mathcal{L} : \Sigma_A \to \Sigma_O$, we do not
  have much motivation to introduce these tuples,
  $\mathopen{<}\cancel{\Sigma_A}, \cancel{\Sigma_O}, \mathcal{L},
  \cancel{S}\mathclose{>}$. Instead of composing ACGs, we compose just
  the lexicons.

On top of that, we will want to define signatures and lexicons to yield
multiple different object languages (at least one for the string
representations of forms and one for the semantic representations of
meaning), which would mean having at least two different ACGs which are
somehow linked together (e.g. by sharing the same abstract
signature). However, since these two grammars are so interrelated, we
would like to reason about them in a way that makes these relationships
explicit and thus easier to study as opposed to considering a collection
of grammars which are somehow implicitly similar.}

\subsection{Patterns of Composition in ACGs}
\label{ssec:acg-patterns}

Before we turn to our proposal, we will quickly recall some of the other
styles of composition used in ACGs (a more detailed exposition can be
found in \cite{pogodalla2012controlling}). Besides making the object
signature of one grammar be the abstract signature of another as we saw
before (Figure~\ref{fig:acg-serial-comp}), we can have two grammars
share the same abstract signature allowing us to transduce between terms
in the two object signatures (Figure~\ref{fig:acg-parallel-comp}).

NB: \emph{Transduction} is a process in which mapping an object $o_1$ to
(potentially many) other objects $o_2$ is realized by first inverting
one function $f_1$ to find the object's antecedents $a$ and then
applying to them another function $f_2$, i.e. an object $o_1$ is
transduced to an object $o_2$ whenever there exists an $a$ such that
$f_1(a) = o_1$ and $f_2(a) = o_2$. In our example, terms from
$\mathcal{O}( \mathopen{<} \Sigma_1, \Sigma_2, \mathcal{L}_{12}, S
\mathclose{>} )$ are transduced to terms from $\mathcal{O}( \mathopen{<}
\Sigma_1, \Sigma_3, \mathcal{L}_{13}, S \mathclose{>} )$ by inverting
the lexicon $\mathcal{L}_{12}$ to find an antecedent in $\mathcal{A}(
\mathopen{<} \Sigma_1, \Sigma_2, \mathcal{L}_{12}, S \mathclose{>} )$
and then applying to it the lexicon $\mathcal{L}_{13}$.

\begin{figure}[t]
  \centering
  \begin{subfigure}[b]{0.4\textwidth}
    \centering
    \includegraphics[height=0.2\textheight]{diagrams/parallel-over-serial.pdf}
    \caption{\label{fig:parallel-over-serial} Transduction from an
      abstract signature.}
  \end{subfigure}
  \qquad
  \begin{subfigure}[b]{0.4\textwidth}
    \centering
    \includegraphics[height=0.2\textheight]{diagrams/serial-over-parallel.pdf}
    \caption{\label{fig:serial-over-parallel} Constraint on an abstract
      signature.}
  \end{subfigure}
  \caption{\label{fig:acg-comp-patterns} Mixing the ACG modes of
    composition.}
\end{figure}

These two modes of composition can be fruitfully mixed. Consider the
graph on Figure~\ref{fig:parallel-over-serial} in which we introduce a
syntactic signature $Syntax$ where different terms are only allowed to
yield the same string whenever they represent different parses of a
syntactically ambiguous phrase. This means that purely semantic
distinctions such as scope ambiguity have to be moved to a more abstract
signature ($SyntSem$) from whose terms the semantic representation can
be derived by a function ($\mathcal{L}_{InterSem}$). This approach is
studied more deeply in \cite{pogodalla2007generalizing}.

Another interesting composition pattern is to take an existing structure
of grammars and add a new abstract signature on top
(Figure~\ref{fig:serial-over-parallel}). This new signature ($Constr$)
can be used to further constrain the items in the previously
abstract-most signature ($Synt$). In \cite{pogodalla2012controlling},
this is what the authors use to develop their constraints on
extraction. Each of their constraints is presented as a new abstract
signature constraining the original syntactic signature.

However, a question that the authors do not pose, but which is of great
interest to us, is how to combine all these constraints. If we were to
apply the same pattern repeatedly to get a structure like the one on
Figure~\ref{fig:stacked-constraints}, we would have to deal with
successively more complex type systems on each level. This is due to the
condition that a lexicon has to be a homomorphism and thus it must,
among other things, always map well-typed terms to well-typed
terms. Because of this, the terms in the abstract language of the
signature we introduce to handle the second constraint must also satisfy
the first constraint, since mapping them by $\mathcal{L}_{Constr_2}$
must yield a well-typed term in the signature corresponding to the first
constraint. In the end, whenever we wish to add a new signature to
handle another constraint, we have to acknowledge and reimplement in it
all of the existing constraints. This effectively renders the pattern
useless since we might just as well consider the final abstract-most
signature handling all the constraints ($Constr_{1 \land 2}$) and attach
it directly to the syntactic signature ($Synt$) by taking the
composition of the lexicons in between ($\mathcal{L}_{Constr_1} \circ
\mathcal{L}_{Constr_2}$).

\begin{figure}[t]
  \centering
  \begin{subfigure}[b]{0.4\textwidth}
    \centering
    \includegraphics[height=0.2\textheight]{diagrams/stacked-constraints.pdf}
    \caption{\label{fig:stacked-constraints} Stacking the constraints
      vertically.}
  \end{subfigure}
  \qquad
  \begin{subfigure}[b]{0.4\textwidth}
    \centering
    \includegraphics[height=0.2\textheight]{diagrams/constraints-side-by-side.pdf}
    \caption{\label{fig:constraints-side-by-side} Laying out the
      constraints side by side.}
  \end{subfigure}
  \caption{Constraining a syntactic signature with multiple
    constraints.}
\end{figure}

What we would like instead is to put the signatures side by side such
that they do not depend on each other
(Figure~\ref{fig:constraints-side-by-side}). However, our language of
syntactic structures is no longer defined by mapping a single abstract
language through a lexicon. It is outside of the scope of the common
formalization of ACGs that we presented in~\ref{ssec:usual-acgs}. How do
we define this language in a manner that is consistent with our current
definitions of ACGs? We can arrive at the answer by examining how the
graphical representations of ACGs are connected to the languages they
define.

\subsection{Interpreting ACG Diagrams}

We will associate a language to every node of an ACG diagram. Before we
begin, we notice that every ACG diagram formed by the above modes of
composition is an arborescence\footnote{An arborescence is a directed
  rooted tree, where every edge points away from the root.} and
therefore every node has at most one predecessor (the parent). To the
root of the tree, the abstract-most signature in the graph, we will
assign the abstract language generated by that signature and its
distinguished type. To all the other nodes, we will assign the result of
mapping their parent's language by the lexicon linking the two. By
homomorphism of the lexicons, we can find distinguished types for all
the non-root nodes. They are simply the images of their parents'
distinguished types by the connecting lexicons. For all of the nodes, it
is true that every term in the language assigned to the node has the
node's distinguished type.

We can confirm that the above graphical model of defining ACGs is
coherent.

First, we can show that our interpretation is sound meaning that any
arborescence adorned with signatures on the nodes, compatible lexicons
on the edges and a distinguished type on the root can only define ACG
languages (abstract or object languages). This is trivial, since the
language corresponding to the root is an ACG abstract language and the
class of ACG languages is closed under transformation by a lexicon and
therefore all of the languages corresponding to the descendants of the
root are ACG languages as well.

Second, we can try and show some kind of completeness. However, there is
not much formalization done to describe a ``system of ACGs that share
some common signatures'' (something that our generalization will
fix). We can nevertheless show that it is possible to express any ACG
using our graphical model. Given a grammar $\mathcal{G} =
\mathopen{<}\Sigma_A, \Sigma_O, \mathcal{L}, S\mathclose{>}$, we just
produce a tree whose root has the signature $\Sigma_A$ and distinguished
type $S$ and which is connected to a single child node having the
signature $\Sigma_O$ by an edge labelled by the lexicon
$\mathcal{L}$. Then the two nodes of our tree define both the abstract
and the object language defined by $\mathcal{G}$. Similarly, we could
argue that the languages we aim to define using ACG diagrams (such as
the ones on Figures \ref{fig:acg-comp-modes} and
\ref{fig:acg-comp-patterns}) are covered by our assignment of languages
to nodes of the diagrams.

We have now given a formal account of how we can read language
definitions from ACG diagrams. Still, we have kept our analysis to the
arborescent diagrams used in current ACG literature, where every node
has at most one parent. We will now proceed to generalize it to cases
where nodes have more than just one parent.

To make the generalization more visible, we will switch from an
algebraic manner of presentation to a relational one. Before, we stated
that the language of a node is the image of its parent's language.

$$
Language(u) = Lexicon(Language(Parent(u)))
$$

Now, we will restate this as saying that the language of a node is the
set of terms that have an antecedent in the parent's language.

$$
Language(u) = \{x \in \Lambda(Signature(u)) \mid \exists a \in
Language(Parent(u)), \; Lexicon(a) = x\}
$$

This definition works for non-root nodes of an ACG diagram. In the root
case, we have to constrain the language by stating that that the terms
of the language must have the root's distinguished type. This is a
property that is also true in the non-root cases due to the homomorphic
properties of lexicons. If we wanted to unify the definitions of the
root language and the non-root languages, we would have to check that a
term of the language has the distinguished type and that it has an
antecedent in the parent, if it has any parent. A relational statement
of this unification could look like this.

\begin{align*}
Language(u) = \{&x \in \Lambda(Signature(u)) \mid Type(x) = DistinguishedType(u) \\
&\land \forall p \in Parents(u), \; \exists a \in Language(p), \; Lexicon_{(p,u)}(a) = x\}
\end{align*}

If the node has no parents, no constraint is enforced. If it does have
some, we enforce the antecedent constraint. This shows us a natural
generalization to the case where nodes can have an arbitrary number of
parents. If we transpose this back to the algebraic style of
presentation, we get the following expression.

$$
Language(u) = DistinguishedlyTyped(u) \quad \cap \bigcap_{p \in Parents(u)} Lexicon_{(p,u)}(Language(p))
$$

\section{Graphical Abstract Categorial Grammars}

In this section, we distill the reasoning about generalizing ACG
diagrams to graphs in a definition of an extension for ACGs. We then
examine its key properties and implications.

We define a \emph{graphical abstract categorial grammar} as a quadruple
$\mathcal{G} = \mathopen{<} G, \Sigma, S, {\mathcal{L}} \mathclose{>}$
where $G$ is a directed graph with vertices $V(G)$ and edges $E(G)$,
$\Sigma$ and $S$ are labelings assigning signatures and distinguished
types, respectively, to the vertices $V(G)$ and ${\mathcal{L}}$ is a
labeling assigning lexicons to the edges $E(G)$. Furthermore, a
well-formed graphical ACG satisfies the following conditions:

\begin{itemize}
\item $G$ is a directed acyclic graph.
\item For all $(u,v) \in E(G)$, $\mathcal{L}_{(u,v)} : \Sigma_u \to
  \Sigma_v$.
\item For all $(u,v) \in E(G)$, $\mathcal{L}_{(u,v)}(S_u) = S_v$.
\end{itemize}

We will say that a node $u$ is \emph{more abstract} than a node $v$
whenever $(u,v)$ belongs to the strict partial order induced by the
edges of $G$ (i.e. there is a path from $u$ to $v$).

We let the nodes in a graphical ACG $\mathcal{G} = \mathopen{<} G,
\Sigma, S, \mathcal{L} \mathclose{>}$ define two kinds of languages. The
\emph{intrinsic languages} $\mathcal{I}_{\mathcal{G}}$, which are
constrained only by the type signature described in the node itself, and
the \emph{extrinsic languages} $\mathcal{E}_{\mathcal{G}}$, which are
also constrained by type signatures in more abstract nodes of the
graph. These two notions correspond to those of abstract languages and
object languages in ACGs.

$$
\mathcal{I}_{\mathcal{G}}(v) = \{t \in \Lambda(\Sigma_v)
\mid\ \vdash_{\Sigma_v} t : S_v\}
$$
$$
\mathcal{E}_{\mathcal{G}}(v) = \mathcal{I}_{\mathcal{G}}(v) \cap
\bigcap_{(u,v) \in E} \mathcal{L}_{(u,v)}(\mathcal{E}_{\mathcal{G}}(u)).
$$

From the above definition, we see that the ACG diagrams we introduced
before are a special case of graphical ACGs where the graph is an
arborescence.

\subsection{On the Expressivity of Graphical ACGs}

\begin{observation}
  The set of intrinsic languages definable by graphical ACGs is the same
  as the set of abstract languages definable by ACGs.
\end{observation}

\begin{theorem}
  The set of extrinsic languages definable by graphical ACGs is the set
  of object languages definable by ACGs closed under intersection and
  transformation by a lexicon.
\end{theorem}

\begin{proof}
  First, we prove that an extrinsic language can be always constructed
  from object languages by intersections and transformations by
  lexicons. Let us consider any graphical ACG $\mathcal{G} =
  \mathopen{<} G, \Sigma, S, \mathcal{L} \mathclose{>}$ and some
  topological ordering $v_1, \ldots, v_n$ of its nodes $V(G)$ (an
  ordering such that more abstract nodes always precede less abstract
  ones). Before we start, we will remark that all the intrinsic
  languages $\mathcal{I}_{\mathcal{G}}(v_i)$ are object languages, since
  intrinsic languages are abstract languages which are in turn just a
  special case of object languages.

  We proceed by induction on the topological ordering:

  \begin{itemize}
    \item For $v_1$, the case is trivial. $v_1$ has no predecessors and
      so $\mathcal{E}_{\mathcal{G}}(v_1) =
      \mathcal{I}_{\mathcal{G}}(v_1)$, which is an object language.
    \item For any other node $v_n$, we look at the definition of
      $\mathcal{E}_{\mathcal{G}}(v_n)$ and remark that the only
      operators are intersection and application of a lexicon with the
      operands being $\mathcal{I}_{\mathcal{G}}(v_n)$ (which is an
      object language) and the extrinsic languages of its predecessors
      (which can be constructed from object languages using
      intersections and lexicons thanks to the induction
      hypothesis). Therefore even $\mathcal{E}_{\mathcal{G}}(v_n)$ can
      be constructed from object languages using only intersections and
      lexicons.
  \end{itemize}

  Now we have to prove the converse. We do so by showing that extrinsic
  languages contain object languages and are closed under intersection
  and transformation by a lexicon.

  \begin{itemize}
    \item Let $L$ be an object language defined by the ACG $\mathcal{G}
      = \mathopen{<} \Sigma_A, \Sigma_O, \mathcal{L},
      S\mathclose{>}$. We take the graph $G$ with $V(G) = \{A, O\}$ and
      $E(G) = \{(A,O)\}$, we label the nodes $A$ and $O$ with $\Sigma_A$
      and $\Sigma_O$ as signatures and $S$ and $\mathcal{L}(S)$ as
      distinguished types respectively. Finally, we label the edge
      $(A,O)$ with $\mathcal{L}$ and we get a graphical ACG $G'$ in
      which $\mathcal{E}_{\mathcal{G'}}(O) = \mathcal{O}(G) = L$.

    \item Let $L_1$ and $L_2$ be two extrinsic languages defined by the
      nodes $u_1$ and $u_2$ in some graphical ACGs $\mathcal{G}_1$ and
      $\mathcal{G}_2$, respectively. We can construct a new graphical
      ACG $\mathcal{G}'$ by taking the union of the two ACGs (union of
      graphs (union of vertices and edges) and unions of the
      labelings).\footnote{Assuming, without loss of generality, that
        the sets of vertices of the two graphs are disjoint.} We will
      then add a new node $v$ such that $\mathcal{E}_{\mathcal{G}'}(v) =
      L_1 \cap L_2$.

      The signature labeling $v$ will be the union of the two signatures
      labeling $u_1$ and $u_2$, ensuring that $\forall u \in \{u_1,
      u_2\}$, $\mathcal{I}_{\mathcal{G}'}(v) \supseteq
      \mathcal{I}_{\mathcal{G}'}(u) \supseteq
      \mathcal{E}_{\mathcal{G}'}(u)$. We can use either of the two
      distinguished types that label $u_1$ or $u_2$ to label $v$ (if
      they are different, then the intersection $L_1 \cap L_2$ is
      trivially empty). Finally, we will add two edges leading from
      $u_1$ and $u_2$ to $v$, both labeled with the identity lexicon. We
      will now show that $\mathcal{E}_{\mathcal{G}'}(v) = L_1 \cap L_2$.

      \begin{align*}
        \mathcal{E}_{\mathcal{G}'}(v) &= \mathcal{I}_{\mathcal{G}'}(v)
        \cap \mathcal{L}_{(u_1,v)}(\mathcal{E}_{\mathcal{G}'}(u_1)) \cap
        \mathcal{L}_{(u_2,v)}(\mathcal{E}_{\mathcal{G}'}(u_2)) \\ &=
        \mathcal{I}_{\mathcal{G}'}(v) \cap
        \mathcal{E}_{\mathcal{G}'}(u_1) \cap
        \mathcal{E}_{\mathcal{G}'}(u_2) \\ &=
        \mathcal{E}_{\mathcal{G}'}(u_1) \cap
        \mathcal{E}_{\mathcal{G}'}(u_2) \\ &=
        \mathcal{E}_{\mathcal{G}}(u_1) \cap
        \mathcal{E}_{\mathcal{G}}(u_2) \\ &= L_1 \cap L_2
      \end{align*}

      The simplifications use the facts that: the two new lexicons are
      identities, that $\mathcal{I}_{\mathcal{G}'}(v) \supseteq
      \mathcal{E}_{\mathcal{G}'}(u_1) \cap
      \mathcal{E}_{\mathcal{G}'}(u_2)$, that
      $\mathcal{E}_{\mathcal{G}'}(u_1) = \mathcal{E}_{\mathcal{G}}(u_1)$
      and $\mathcal{E}_{\mathcal{G}'}(u_2) =
      \mathcal{E}_{\mathcal{G}}(u_2)$ and finally that
      $\mathcal{E}_{\mathcal{G}}(u_1) = L_1$ and
      $\mathcal{E}_{\mathcal{G}}(u_2) = L_2$.

  \item Let $L \subseteq \Lambda(\Sigma_O)$ be a language which is the
    result of mapping an extrinsic language $L' \subseteq
    \Lambda(\Sigma_A)$ through the lexicon $\mathcal{L} : \Sigma_A \to
    \Sigma_O$. We will show that $L$ is also an extrinsic language. Let
    $u$ be the node that defines the extrinsic language $L$ in some
    graphical ACG $\mathcal{G}$. We construct a new graphical ACG
    $\mathcal{G}'$ by adding a new vertex $v$ to $\mathcal{G}$ labeled
    with the signature $\Sigma_O$ and the distinguished type
    $\mathcal{L}(S)$ where $S$ is the distinguished type labeling
    $u$. To this graph, we conjoin the edge $(u,v)$ and label it with
    $\mathcal{L}$. Now we verify that $\mathcal{E}_{\mathcal{G}'}(v) =
    L$.

    \begin{align*}
      \mathcal{E}_{\mathcal{G}'}(v) &= \mathcal{I}_{\mathcal{G}'}(v) \cap
      \mathcal{L}(\mathcal{E}_{\mathcal{G}'}(u)) \\
      &= \mathcal{L}(\mathcal{E}_{\mathcal{G}'}(u)) \\
      &= \mathcal{L}(\mathcal{E}_{\mathcal{G}}(u)) \\
      &= \mathcal{L}(L') \\
      &= L
    \end{align*}

    The assumptions used here are in turn: that
    $\mathcal{I}_{\mathcal{G}'}(v) \supseteq
    \mathcal{L}(\mathcal{E}_{\mathcal{G}'}(u))$, that
    $\mathcal{E}_{\mathcal{G}'}(u) = \mathcal{E}_{\mathcal{G}}(u)$, that
    $\mathcal{E}_{\mathcal{G}}(u) = L'$ and that $\mathcal{L}(L') = L$.
  \end{itemize}
\end{proof}

\begin{corollary}
  The set of extrinsic languages definable by graphical ACGs is the same
  as the set of object languages definable by ACGs if and only if the
  set of object languages definable by ACGs is closed under
  intersection.
\end{corollary}

We conclude this subsection with a few words to motivate the need we
feel for this (possibly) more expressive formalism. We believe that in
order to manage the development of wide-coverage grammars, the grammars
need to be kept as simple as possible, which means, among other things,
that independent constraints should be implemented separately and not
complected together (we saw the difficulties caused by this already on a
microscopic fragment in section~\ref{sec:together}). This means that if
we were to define, e.g., a partial grammar in which we would only
enforce agreement and another grammar in which we would only enforce the
proper use of relative pronouns, we would like the language in which
both agreement and proper use of relative pronouns are respected to be
expressible in our formalism, and on top of that, we would like this
language to be easily expressed in terms of the two grammars for the
individual constraints. We now have a requirement on the formalism in
which we would like to use: we want a formalism where the set of all the
``useful'' definable languages is closed on intersection. If the object
languages of ACGs are indeed closed on intersection, then our extension
is just a convenience for expressing intersection more easily. However,
if they are not, we believe our extension adds a critical level of
expressivity needed for simple and modular wide-coverage grammars.

\subsection{On the Decidability of Graphical ACGs}
\label{ssec:graphical-decidability}

We turn to the problem of parsing with graphical ACGs. Parsing with ACGs
is a problem on the limits of decidability. When we consider ACGs whose
abstract signatures only contain types of order 2 or less, parsing is
possible in polynomial type for both linear \cite{salvati2005problemes}
and non-linear \cite{kanazawa2007parsing} type systems (such as the one
we introduced in~\ref{sec:acg}). If the order of the abstract signature
exceeds 2, the membership problem for linear ACGs becomes decidable if
and only if the satisfiability problem in the multiplicative exponential
fragment of linear logic is decidable \cite{de2004vector}, which is an
open problem. On the other hand, by extending the type system to include
dependent types, parsing can become undecidable too since in such a
system type checking itself is undecidable.

Our intent here is to show that by extending the formalism we have not
made the problem of parsing more difficult, at least in terms of
decidability if not in terms of complexity. Let us therefore imagine a
graphical ACG where every edge (two signatures, distinguished types and
a lexicon) is an ACG for which there is an effective parsing algorithm
and type checking is decidable in all the signatures of the graphical
ACG. Then we can prove the existence of a simple effective parsing
algorithm for any of the extrinsic languages defined in the graphical
ACG.

We will be testing whether the term $o$ belongs to
$\mathcal{E}_{\mathcal{G}}(v)$. Our algorithm first checks whether $o$
belongs to $\mathcal{I}_{\mathcal{G}}(v)$ by checking its type. If it
does, it proceeds to verify that it is also in
$\mathcal{E}_{\mathcal{G}}(v)$ by finding antecedents in all of its
abstract predecessors $u_i$, if there are any. We can use the
presupposed parsing algorithm for the ACG $\mathopen{<} \Sigma_{u_i},
\Sigma_v, \mathcal{L}_{(u_i,v)}, S_{u_i}\mathclose{>}$ to find all the
abstract terms $\alpha_{i,j}$ having type $S_{u_i}$ and being
antecedents to $o$ w.r.t. the lexicon $\mathcal{L}_{(u_i,v)}$. Such
terms belong to $\mathcal{I}_{\mathcal{G}}(u_i)$, but in order to prove
that $o$ is in $\mathcal{E}_{\mathcal{G}}(v)$, we need to find an
antecedent that belongs to $\mathcal{E}_{\mathcal{G}}(u_i)$. What we do
is we recursively apply our parsing algorithm to all of the
antecedent-candidates $\{\alpha_{i,j} \mid \forall j\}$ until we find
one which belongs $\mathcal{E}_{\mathcal{G}}(u_i)$. Since the graph is
acyclic, we are guaranteed that this recursive traversal will
terminate. We repeat this for all the predecessors $u_i$ and if we
succeed to find antecedents in all of their extrinsic languages, we have
confirmed the membership of $o$ in $\mathcal{E}_{\mathcal{G}}(v)$ and
also built its abstract antecedent(s).

\subsection{On Grammar Engineering with Graphical ACGs}
\label{ssec:graphical-engineering}

Having covered some of the formal properties of graphical ACGs, we will
get back to building grammars. Armed with this formalism, we can solve
the puzzle of incorporating the three constraints of sections
\ref{sec:negation}, \ref{sec:extraction} and \ref{sec:agreement}. Our
graph $G$ will consist of 6 nodes, $V(G) = \{Neg, Ext, Agr, Synt, Str,
Sem\}$ with edges going from $Neg$, $Ext$ and $Agr$ to $Synt$ and from
$Synt$ to $Str$ and $Sem$ (see Figure~\ref{fig:puzzle-solution}). The
signatures and distinguished types for the first three vertices are
exactly those that have been presented in the first three sections of
this chapter. The lexicons for the edges connecting these constraint
nodes to $Synt$ all behave according to the scheme below:

\begin{align*}
  \mathcal{L}_{(constraint,Synt)}(X_{constant_{index}}) &= C_{constant} \\
  \mathcal{L}_{(constraint,Synt)}(TYPE\_FEATURES)  &= TYPE
\end{align*}

The $\{Synt, Str, Sem\}$ subgraph is identical to the pair of ACGs
described in \ref{sec:acg} except for some additions that need to be
made so that the fragment covers the wordforms that we have introduced
in the course of discussing the three constraints (prepositions,
relative pronouns, the negative particle $ne$\ldots). However, the types
of these new constants will not be tied to any of those constraints.

\begin{align*}
  C_{de} &: NP \limp N \limp N \\
  C_{qui} &: (NP \limp S) \limp N \limp N \\
  C_{dit\ que} &: NP \limp S \limp S \\
  C_{dorment} &: NP \limp S \\
  C_{ne_{tv}} &: (NP \limp NP \limp S) \limp (NP \limp NP \limp S) \\
  C_{ne_{iv}} &: (NP \limp S) \limp (NP \limp S) \\
  C_{ne_{cv}} &: (NP \limp S \limp S) \limp (NP \limp S \limp S)
\end{align*}

This approach lets us describe the high-level combinatorics of language
that are of interest to the categorial grammarian in a manner which is
idiomatic in academic literature (using the simple and focused types
$N$, $NP$ and $S$). At the same time, we can actually use this textbook
grammar as a basis for a more realistic grammar which handles
constraints such as agreement and extraction islands. Furthermore, these
additional constraints are also defined in a manner that has its
precedent in existing research. When the authors of
\cite{pogodalla2012controlling} decided to put the constraints of
extraction under a microscope and try to encode them in ACGs, they
defined their constraints in exactly the same way as our architecture
does. This graphical architecture lets us write the syntactic kernel and
the additional constraints independently and using two different styles
that have both been shown useful by their own merits, which we see as
indicative of enabling good modularity.

\begin{figure}[t]
  \centering
  \begin{subfigure}[b]{0.4\textwidth}
    \centering
    \includegraphics[height=0.2\textheight]{diagrams/puzzle-solution.pdf}
    \caption{{\label{fig:puzzle-solution} The graph of a grammar
        enforcing all of the constraints discussed in this chapter.}}
  \end{subfigure}
  \qquad
  \begin{subfigure}[b]{0.4\textwidth}
    \centering
    \includegraphics[height=0.2\textheight]{diagrams/both-patterns.pdf}
    \caption{{\label{fig:both-patterns} Combining the two ACG patterns
        from~\ref{ssec:acg-patterns} in a single graphical ACG.}}
  \end{subfigure}
  \caption{Practical architectures for graphical ACGs.}
\end{figure}

In the above approach, we have defined a language by writing a grammar
which provides us with useful structures but which does not enforce
complete correctness. We have then refined this grammar by tacking on
constraints ``on the side''. This approach is generally useful and can
be seen, e.g., in \cite{van2004concepts} to define the formal
programming language kernel on which a formal semantics is
established. In our case, we are interested in having intelligible
syntactic structures too as we would like to compute their semantic
representations. The syntax-semantics interface already presents a
significant challenge and having less-than-ideal syntactic structures to
start with does not make formalizing it any easier.

We will show one more persuasive scenario whose goal is to show the
importance of graphical ACGs for having a clean syntax-semantics
interface. In \ref{ssec:acg-patterns}, we have seen an ACG pattern that
put semantic ambiguities into an abstract signature that is higher than
syntax. But there was also another pattern, that of adding a new
abstract signature to constrain syntactic terms. We can combine both of
them in graphical ACGs by superposing their two graphs to get the
structure we can see on Figure~\ref{fig:both-patterns}.

If we try to reap the benefits of both patterns and try to replicate the
same in classical ACGs (arborescent graphs), we hit a snag: we cannot
have both the $Constr$ and $SyntSem$ nodes as the direct predecessors of
the $Syntax$ node. We can make $Constr$ be the predecessor of $Syntax$,
but then $SyntSem$ will be connecting $Constr$ and $Sem$ and the
syntax-semantics interface will have to deal with the types which were
needed to implement syntactic constraints and have no place in a
syntax-semantics interface. If we try it the other way, it is not much
better. $Constr$ will end up having to work with the more complicated
type system of $SyntSem$ instead of being able to talk only about the
$Syntax$ it is interested in.

By putting the constraining signature in a part of the graph unreachable
from the syntax-semantics transducer, we can shield the syntax-semantics
interface from the gnarly implementation details of the rules of
syntax. This in turn gives us a guideline as to what information should
be expressed in the syntactic signature itself and which should be
hidden in constraints: the syntactic signature should only contain
information relevant to the syntax-semantics interface.

This might then raise a question whether the chosen syntactic signature
will end up being suitable for expressing all of the syntactic
constraints. However, this can be resolved by deepening the constraint
subgraph and adding a ``compatibility layer'' in the form of a new
signature which provides different syntactic structures and maps them to
the canonical ones. Constraints can then opt in to controlling the
syntactic structures offered by this new signature.

\subsection{On Alternative Interpretations of Graphical ACGs}

Our definitions of intrinsic and extrinsic languages have consequences
which might be counter-intuitive and not correspond to the grammarian's
intent. We will describe what these consequences are and propose an
alternative notion of a graphical ACG language which aims to reconcile
them.

The first phenomenon that deserves pointing out occurs whenever two
paths from nodes $u_1$ and $u_2$ converge on a single node $v$. If we
then take a term $\alpha$ from $\mathcal{E}_{\mathcal{G}}(u_1)$, the
term $\beta = \mathcal{L}_{(u_1,v)}(\alpha)$ belongs to
$\mathcal{I}_{\mathcal{G}}(v)$ but not necessarily to
$\mathcal{E}_{\mathcal{G}}(v)$ since it might be the case that there is
no antecedent for $\beta$ in $\mathcal{E}_{\mathcal{G}}(u_2)$. In other
words, mapping a member of $\mathcal{E}_{\mathcal{G}}(u)$ using the
lexicon $\mathcal{L}_{(u,v)}$ does not guarantee that the result belongs
to $\mathcal{E}_{\mathcal{G}}(v)$, which is a different behavior from
that of object languages in ACGs (or that of extrinsic languages in
arborescent graphical ACGs).

This highlights some limitations of our notion of an extrinsic
language. Consider once more the graphical ACG of
Figure~\ref{fig:both-patterns}. We have more than one path converging to
$Syntax$. We will consider a term $\alpha$ of
$\mathcal{E}_{\mathcal{G}}(Sem)$. $\alpha$ has an antecedent in
$\mathcal{E}_{\mathcal{G}}(SyntSem)$ which has a descendant $\beta$ in
$\mathcal{I}_{\mathcal{G}}(Syntax)$. However, because of the constraint
$Constr$ on $Syntax$, $\beta$ can be a syntactic term outside of
$\mathcal{E}_{\mathcal{G}}(Syntax)$ and yielding an ungrammatical
sentence. The notion of an extrinsic language is insufficient to express
the set of meanings from $\Lambda(\Sigma_{Sem})$ which are expressible
only by grammatical sentences. Defining languages like this would be of
particular interest in cases like the one in
Figure~\ref{fig:transducer-grammar} where we have a chain of signatures
representing the different levels of linguistic description whose terms
are in many-to-many relationships expressed through transducers. In such
a scenario, it would be desirable to be able to express the set of
descriptions belonging to one level, e.g. morphology, that correspond to
valid descriptions on all the other levels.

\begin{figure}[t]
  \centering
  \begin{subfigure}[b]{0.4\textwidth}
    \centering
    \includegraphics[width=\textwidth]{diagrams/transducer-grammar.pdf}
    \caption{{\label{fig:transducer-grammar} A graphical ACG connecting
        (constrained) levels of linguistic description through
        transducers.}}
  \end{subfigure}
  \qquad
  \begin{subfigure}[b]{0.4\textwidth}
    \centering
    \includegraphics[height=0.2\textheight]{diagrams/diamond-grammar.pdf}
    \caption{{\label{fig:diamond-grammar} A graphical ACG demonstrating
        two different paths between two nodes.}}
  \end{subfigure}
  \caption{Graphical ACGs posing challenges to extrinsic languages.}
\end{figure}

Compared to ACG object languages (or to extrinsic languages of
arborescent graphical ACGs), we lose another guarantee. In
arborescences, the central role of the root abstract-most signature
guarantees us that whenever a term belongs to an extrinsic language, we
can find terms in all the other extrinsic languages that share the same
abstract-most term. On the other hand, in non-arborescent graphical
ACGs, having a term in an extrinsic language of some node $v$ only
guarantees the presence of corresponding terms in the extrinsic
languages of its ancestors and the intrinsic languages of their
descendants.

The second puzzling scenario is the diamond shape on
Figure~\ref{fig:diamond-grammar}. If we consider a term $\delta$ of
$\mathcal{E}_{\mathcal{G}}(Bottom)$, we know that there exist
antecedents $\beta$ and $\gamma$ in $\mathcal{E}_{\mathcal{G}}(Left)$
and $\mathcal{E}_{\mathcal{G}}(Right)$. And while there definitely exist
antecedents $\alpha$ and $\alpha'$ to both $\beta$ and $\gamma$ in
$\mathcal{E}_{\mathcal{G}}(Top)$, these antecedents can be different,
meaning there does not have to be a single $\alpha \in
\mathcal{E}_{\mathcal{G}}(Top)$ which yields $\beta \in
\mathcal{E}_{\mathcal{G}}(Left)$ and $\gamma \in
\mathcal{E}_{\mathcal{G}}(Right)$ such that they both yield
$\delta$. This means that the result of parsing a term in some
$\mathcal{E}_{\mathcal{G}}(v)$ is a collection of abstract terms with
one term not just per every node $u$ that is more abstract than $v$ but
per every different path leading to $v$. Parsing a term $\delta \in
\mathcal{E}_{\mathcal{G}}(Bottom)$ in our diamond grammar would thus
produce four antecedents $\beta$, $\gamma$, $\alpha$ and $\alpha'$
corresponding to the four paths $[Left,Bottom]$, $[Right,Bottom]$,
$[Top,Left,Bottom]$ and $[Top,Right,Bottom]$, respectively. We contrast
this again with the case of classical ACGs where the result of parsing
can always be represented by associating a term to every node of the
arborescence.\footnote{One could also say that the result of parsing in
  an ACG is just the antecedent in the abstract-most signature and that
  the other terms can then be easily computed by application of
  lexicons. Similarly, we could say that the result of parsing in a
  general graphical ACG is a collection of abstract terms for every
  \emph{maximal} path leading to the node defining the language in
  question. What might be more intuitive, though, is to have some theory
  of graphical ACGs where the result of parsing is a collection of
  antecedents in all the abstract-most nodes.}

We will propose a new way of assigning languages to nodes in a graphical
ACG that will conserve the intuitive properties of ACGs we talked about
above. We do not need to modify the definition of a graphical ACG in
order to do so since the definition is only concerned with the structure
of the graph and its decoration with ACG paraphernalia, not the
languages defined. We will define a third kind of languages based on
graphical ACGs.

Let $\mathcal{G} = \mathopen{<} G, \Sigma, S, \mathcal{L} \mathclose{>}$
be a graphical ACG. We define the \emph{pangraphical language} of node
$u$, $\mathcal{P}_{\mathcal{G}}(u)$, to be the set of terms $t$ such
that there exists a labeling of nodes with terms $T$ satisfying the
following conditions:

\begin{itemize}
  \item $T_u = t$.
  \item For all $v \in V(G)$, $\vdash_{\Sigma_v} T_v : S_v$.
  \item For all $(v,w) \in E(G)$, $\mathcal{L}_{(v,w)}(T_v) = T_w$.
\end{itemize}

It is easy to see that the newly introduced pangraphical languages solve
all of the counter-intuitive quirks of object languages that we
discussed in this subsection.

\begin{itemize}
\item Applying the lexicon $\mathcal{L}_{(u,v)}$ to an element of
  $\mathcal{P}_{\mathcal{G}}(u)$ always yields an element of
  $\mathcal{P}_{\mathcal{G}}(v)$, since the labeling of nodes with terms
  $T$ that proved the former term's membership in the former language
  also proves the latter term's membership in the latter language.

  This definition allows us to directly establish the language of
  meanings which are expressible only by grammatical sentences in
  Figure~\ref{fig:both-patterns} and the language of morphological terms
  that have valid corresponding representations on all the other levels
  of linguistic description in Figure~\ref{fig:transducer-grammar}.
\item The diamond grammar of Figure~\ref{fig:diamond-grammar} is also no
  longer problematic since the labeling of nodes with terms guarantees
  that there is always a single term per node that can generate the term
  being parsed.
\end{itemize}

In the previous section, we took a stab at formalizing ACG diagrams and
finding the simplest extension capable of handling non-arborescent
graphs. The notion of extrinsic languages that we arrived at allowed us
to define a nicely delimited set of languages, the object languages of
ACGs with intersection. Discrepancies in the behavior of extrinsic
languages in arborescent and non-arborescent graphs led us to
pangraphical languages, which behave more in line with our
intuitions. However, the expressivity of pangraphical languages,
especially w.r.t. to extrinsic languages, still remains to be
explored. Due to limitations of both space and time, we will cover the
interesting properties of pangraphical languages only briefly and
partially here.

\begin{description}
  \item[Extrinsic languages can be defined similarly to pangraphical
    languages] \hfill \\

    In the definition of $\mathcal{P}_{\mathcal{G}}(u)$, we replace the
    labeling of nodes with terms by a labeling of all the paths leading
    to $u$.

  \item[$\mathcal{P}_{\mathcal{G}}(u) \subseteq
    \mathcal{E}_{\mathcal{G}}(u) \subseteq
    \mathcal{I}_{\mathcal{G}}(u)$] \hfill \\

    Given a labeling of nodes with terms such that it satisfies the
    conditions in the definition of a pangraphical language, we can show
    by induction on the topological ordering of the graph that every
    term assigned by the labeling belongs to the extrinsic language of
    the same node. Therefore, $\alpha \in \mathcal{P}_{\mathcal{G}}(u)
    \implies \alpha \in \mathcal{E}_{\mathcal{G}}(u)$.

    $\mathcal{E}_{\mathcal{G}}(u) \subseteq
    \mathcal{I}_{\mathcal{G}}(u)$ comes straight from the definition of
    $\mathcal{E}_{\mathcal{G}}(u)$.

    We have thus came up with a series of successively more constrained
    languages: intrinsic languages are constrained by one node,
    extrinsic languages are also constrained by its ancestors and
    pangraphical languages are constrained by the entire graph.

  \item[On arborescences, $\mathcal{P}_{\mathcal{G}}(u) =
    \mathcal{E}_{\mathcal{G}}(u)$.] \hfill \\

    On an arborescence, being a member of an extrinsic language means
    there exists an antecedent in the abstract-most language of the
    arborescence. By virtue of the one-parent property of arborescences,
    we can simply apply all the lexicons in the arborescence to this
    abstract-most antecedent to find a labeling of nodes with terms
    which proves that the member of the extrinsic language is also a
    member of the pangraphical language.

    This confirms that both extrinsic languages and pangraphical
    languages are reasonable generalizations of ACG diagrams since their
    behaviors on arborescences are consistent with each other and with
    that of ACG object languages.

  \item[The set of extrinsic languages is included in the set of
    pangraphical languages.] \hfill \\

    Let us have some node $u$ in a graphical ACG $\mathcal{G}$. We build
    a graphical ACG $\mathcal{G}'$ where every node corresponds to a
    path in $\mathcal{G}$ which ends at $u$ and two nodes are connected
    with an edge whenever one path is an immediate extension of the
    other. We then have $\mathcal{P}_{\mathcal{G'}}([u]) =
    \mathcal{E}_{\mathcal{G}}(u)$.

  \item[Is the set of pangraphical languages included in the set of
    extrinsic languages?] \hfill \\

    That is an interesting question indeed, whose resolution is left as
    an exercise for the reader.

  \item[Pangraphical languages are not ``less decidable'' than object
    languages.] \hfill \\

    Presuming that type checking is decidable for all the signatures of
    the graphical ACGs and that we can parse effectively for all the ACG
    edges in the graph, we can reuse the parsing algorithm we introduced
    in \ref{ssec:graphical-decidability}. We apply this algorithm to
    enumerate all the possible extrinsic parses in a language
    $\mathcal{E}_{\mathcal{G}}(v)$ and keep only those where for each
    ancestor $u$ we have the same abstract term for all paths from $u$
    to $v$.
\end{description}

\subsection{On the Practical Consequences of Graphical ACGs}

In this final subsection, we will enumerate some of the practical
benefits and challenges of using graphical ACGs for implementing
grammar-based systems.

\begin{itemize}
\item The lexicons that go from our constraint signatures ($Neg$, $Ext$
  and $Agr$ in subsection \ref{ssec:graphical-engineering}) are all
  trivial in the sense they map constants to constants. Finding an
  antecedent to a term is then just a question of finding an antecedent
  for every constant in the term (i.e. ``tagging'' the constants of the
  term with antecedent constants). This means that there is an effective
  parsing algorithm for verifying the constraints. Furthermore, this
  algorithm could also be made efficient by introducing filtering
  techniques \cite{guillaume2008toolchain} and statistical supertagging
  \cite{moot2012categorial}.

\item Constraints can introduce type extensions in their abstract
  signatures and the added complexity of these extensions is then
  contained within the constraint. Furthermore, the constraint signature
  can be treated as a ``black box'' that verifies a constraint and in
  the context of an implementation, it can be replaced with a
  special-purpose hand-written procedure that does the same more
  efficiently.

\item The distributed nature of graphical ACGs has its own practical
  applications. Parsing can be made more robust by being able to parse
  sentences which violate some of the constraints. Parsing an
  ungrammatical sentence and identifying the constraints it violates can
  also serve as a basis for a syntax checker.

\item If we decide to define all of our syntactic constraints in
  separate signatures, we will end up with a lot of signatures which
  look exactly like the syntactic signature but with some small
  refinements (see the signatures in sections \ref{sec:negation},
  \ref{sec:extraction} and \ref{sec:agreement}). Thus we will desire
  some concise way of formulating these refinements so that we will not
  need to replicate the entire syntactic signature for every
  constraint.

  We hope that the class of useful refinements is small and that these
  constraint signatures could be systematically constructed from the
  syntactic signature by applying a series of refining
  transformations. This is an interesting engineering challenge awaiting
  anyone willing to design a truly wide-coverage graphical ACG using
  constraints.
\end{itemize}