### 2013-06-16 thesis.tex

\documentclass{report}

%% \usepackage{sdrt}
\usepackage{tikz}
\usepackage{amsmath}
\usepackage{amsthm}
\usepackage{xcolor}
\usepackage{graphicx}
\usepackage{gb4e}
\usepackage{bussproofs}
\usepackage{hyperref}
\usepackage{subcaption}
\usepackage{cancel}
\usepackage[yyyymmdd,hhmmss]{datetime}

%% Linear implication from Alessio Guglielmi
%% https://groups.google.com/forum/?fromgroups=#!topic/comp.text.tex/0B4C3F_BsVI
\def\limp {\mathbin{{-}\mkern-3.5mu{\circ}}}

\newtheorem*{observation}{Observation}
\newtheorem*{theorem}{Theorem}
\newtheorem*{corollary}{Corollary}

\begin{document}

\author{Ji\v{r}\'{i} Mar\v{s}\'{i}k}
\title{Master Thesis}
\maketitle

\currenttime

\tableofcontents

\chapter{Preliminaries}

\input{motivation}

\input{acg}

\input{ig}

\input{my-work-rewritten}

\input{the-last-part}

\bibliographystyle{plain}
\bibliography{biblio}

\end{document}