### 2013-06-17 outline.tex

\section{Outline}

In the rest of this chapter, we proceed to introduce two grammatical
formalisms that are of interest to our work, the formalism of abstract
categorial grammars and the formalism interaction grammars. We talk
about a formal connection between the two and its value with respect to
constructing our grammar.

In Chapter~\ref{chap:implementation}, we present our software
implementation of the ACG machinery which will enable us to write and
test grammars lexicalized by a lexicon.

In Chapter~\ref{chap:constraints}, we take the first steps into
designing a wide-coverage grammar by showing the treatment of several
disparate linguistic constraints in the framework of ACGs and the
challenges inherent in trying to combine them all in a single grammar.

Chapter~\ref{chap:gacg} builds on the problems highlighted in
Chapter~\ref{chap:constraints} and introduces a generalization of
abstract categorial grammars that is geared towards solving these
problems.

We conclude our findings with Chapter~\ref{chap:conclusion}, in which we
summarize our findings and point out potential directions for future
work.