### 2013-06-17 thesis.tex

\documentclass{report}

\usepackage{amsmath}
\usepackage{amsthm}
\usepackage{xcolor}
\usepackage{graphicx}
\usepackage{gb4e}
\usepackage{bussproofs}
\usepackage{hyperref}
\usepackage[labelformat=simple]{subcaption}
\renewcommand\thesubfigure{(\alph{subfigure})}
\usepackage{cancel}

%% Linear implication from Alessio Guglielmi
%% https://groups.google.com/forum/?fromgroups=#!topic/comp.text.tex/0B4C3F_BsVI
\def\limp {\mathbin{{-}\mkern-3.5mu{\circ}}}

\newtheorem*{observation}{Observation}
\newtheorem*{theorem}{Theorem}
\newtheorem*{corollary}{Corollary}

\input{front-page}

\input{abstract}

\tableofcontents

\chapter{Preliminaries}
\input{motivation}
\input{outline}
\input{acg}
\input{ig}

\input{implementation}

\input{constraints}

\input{gacg}

\input{conclusion}

\bibliographystyle{plain}
\bibliography{biblio}

\end{document}