### 2013-06-17 acknowledgments.tex
This is the printed version.

\vspace{2\baselineskip}

\topskip0pt
\vspace*{\fill}

\centerline{\textbf{Acknowledgments}}

\vspace{2\baselineskip}

I would like to thank my friends in Nancy for making my year at
Universit\'{e} de Lorraine a wonderful one.

\vspace{\baselineskip}

I would like to thank my supervisor for his enlightening guidance and
his dedication and attention to our work.

\vspace{\baselineskip}

I would like to thank my parents for always supporting me in the things
I have chosen to do.

\vspace{\baselineskip}

\vspace{\baselineskip}

I would like to thank Wes Anderson for \emph{Moonrise Kingdom}.

\vspace*{\fill}