### 2013-06-26 front-page.tex
Fixed links by omitting PDFjam from the pipeline

\documentclass{report}

\usepackage{graphicx}

\begin{document}

\begin{titlepage}
\begin{center}

\newcommand{\HRule}{\includegraphics[width=\textwidth]{images/hrule.png}}

\includegraphics[width=0.9\textwidth]{images/UL_UFR_MI.jpg}\\[1.5cm]

\textsc{\LARGE Master Sciences Cognitives et Applications\\[0.5cm]Sp\'ecialit\'e Traitement Automatique du Langage}\\[1.5cm]

\textsc{\Large Master Thesis}\\[0.65cm]

\HRule \\[0.6cm]
{ \huge \bfseries Towards a Wide-Coverage Grammar: Graphical Abstract Categorial Grammars }\\[0.4cm]

\HRule \\[0.5cm]

June 25, 2013 \\[0.75cm]

\begin{minipage}{0.4\textwidth}
\begin{flushleft} \large
\emph{Author:}\\
Ji\v{r}\'{i} \textsc{Mar\v{s}\'{i}k}\\
~
\end{flushleft}
\end{minipage}
\begin{minipage}{0.4\textwidth}
\begin{flushright} \large
\emph{Supervisor:} \\
Maxime \textsc{Amblard} \\
~
\end{flushright}
\end{minipage}
~\\[0.6cm]
\includegraphics[width=0.35\textwidth]{images/logoLORIA.jpg}~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~\includegraphics[height=0.10\textheight]{images/Calligramme.png}

\end{center}
\end{titlepage}

%\maketitle

%% \newpage
%% ~~ \pagenumbering{gobble}
%% \newpage
%% \pagenumbering{arabic}
%% \setcounter{page}{3}

\end{document}