### 2013-07-02 abstract.tex
Typo in abstract

\begin{abstract}
We present work whose ultimate goal is the creation of a wide-coverage
abstract categorial grammar (ACG) that could be used to automatically
build discourse-level representations. In our work, we advance towards
that goal by laying down the foundations necessary for building
wide-coverage ACGs.

We first examine existing language resources, in particular the Frigram
interaction grammar and its lexicon Frilex, and assess their utility to
building a wide-coverage ACG. We then present our implementation of the
ACG machinery which allows us to experiment with grammars lexicalized by
Frilex. Finally, we consider the challenge of integrating the treatment
of disparate linguistic constraints in a single ACG and propose a
generalization of the formalism: graphical abstract categorial
grammars. The report concludes with an exploration of some of the formal
properties of graphical ACGs.
\end{abstract}