### 2013-07-04 lenls/abstract.tex
Framework for LENLS abstract

\documentclass{article}

\usepackage[utf8]{inputenc}
\usepackage{amsmath}
\usepackage{amsthm}
\usepackage{xcolor}
\usepackage{graphicx}
\usepackage{gb4e}
\usepackage{bussproofs}
\usepackage{hyperref}
\usepackage[labelformat=simple]{subcaption}
\renewcommand\thesubfigure{(\alph{subfigure})}
\usepackage{cancel}
\usepackage[nottoc,numbib]{tocbibind}
\usepackage{pdfpages}
\usepackage{enumitem}

%% Linear implication from Alessio Guglielmi
%% https://groups.google.com/forum/?fromgroups=#!topic/comp.text.tex/0B4C3F_BsVI
\def\limp {\mathbin{{-}\mkern-3.5mu{\circ}}}

\newtheorem*{observation}{Observation}
\newtheorem*{theorem}{Theorem}
\newtheorem*{corollary}{Corollary}

\begin{document}

\title{Towards a Wide-Coverage Grammar: \\ Graphical Abstract Categorial Grammars}
\author{Jiří Maršík \and Maxime Amblard}
\maketitle

\bibliographystyle{plain}
\bibliography{../biblio}

\end{document}