### 2013-12-25 lenls/slides.tex
LENLS slides, backlog commit

\documentclass{beamer}

\usepackage[utf8]{inputenc}
\usepackage{xcolor}
\usepackage{pgfpages}

\definecolor{darkgreen}{rgb}{0., 0.6, 0.}
\definecolor{darkgold}{rgb}{8., 0.6, 0.}
\definecolor{limegreen}{RGB}{175, 255, 175}
\definecolor{limeblue}{RGB}{200, 200, 235}

\hypersetup{pdfstartview={Fit}}
\def\limp {\mathbin{{-}\mkern-3.5mu{\circ}}}

\AtBeginSection[]
{
  \begin{frame}
    \frametitle{Outline}
    \tableofcontents[currentsection]
    \addtocounter{framenumber}{-1}
  \end{frame}
}

\setbeamertemplate{navigation symbols}{}
\setbeamertemplate{footline}
  {\hfill {\normalsize \insertframenumber{}/\inserttotalframenumber{}}}

\setbeameroption{show notes on second screen=right}
%\setbeamertemplate{note page}[plain]

\begin{document}

\title[G-ACGs]{Integration of Multiple Constraints in ACG}
\subtitle{Graphical Abstract Categorial Grammars}
%
\author{Jiří Maršík \and Maxime Amblard}
%
\institute{Université de Lorraine, Laboratoire lorrain de recherche en informatique et ses applications, UMR 7503, Vandoeuvre-lès-Nancy, 54500, France\\
INRIA, Villers-lès-Nancy, 54600, France\\
CNRS, Loria, Vandoeuvre-lès-Nancy, UMR 7503, 54500, France\\
$\,$  \\
\texttt{\{jiri.marsik, maxime.amblard\}@loria.fr}}

\date[October 2013]{October 28, 2013}

\frame{\titlepage \setcounter{framenumber}{1}}

\begin{frame}
\frametitle{Outline}
\tableofcontents
\end{frame}

\section{Introduction}

\begin{frame}
  \frametitle{Introduction}

  \begin{itemize}
    \item Long-term goal: large-scale grammar usable for discourse \vfill
    \item Abstract Categorial Grammars: a tool for discourse and dynamic
      semantics \vfill
    \item Refining grammars to enforce numerous constraints \\ $\Rightarrow$
      complexity \vfill
    \item Extension of ACGs for independent constraints
  \end{itemize}
\end{frame}

\note[itemize]{\item In our team, we are interested in formal treatments of
  discourse-level phenomena.

\item Our tool of choice is the formalism of Abstract Categorial
  Grammars. ACGs provide us with interesting tools to handle discourse and
  dynamic semantics with continuations allowing us to capture dynamic meaning
  and the generality of the lambda calculus allowing us to introduce our own
  constructs such as discourse structures.

\item However, during our work when trying to use ACGs, we observed that it
  proved difficult to try to enforce multiple independent constraints within a
  single grammar without the grammar becoming too complex\ldots

\item \ldots and it is this inadequacy that prompted the extension of the
  formalism that is the focus of our contribution.

\item 0:34}

\section{ACG Preliminaries}

\note{I will now give a short introduction to the ACG formalism.}

\newcommand{\synt}[1]{C_{\textrm{#1}}}

\begin{frame}
  \begin{block}{Abstract languages}
    Defined by a \textcolor{red}{signature} and a
    \textcolor{darkgreen}{distinguished type}.

    \begin{columns}[t]
      \begin{column}{0.5\textwidth}
        \begin{align*}
          \synt{tatous} &: \textcolor{red}{N} \\
          \synt{Laura} &: \textcolor{red}{NP}
        \end{align*}
      \end{column}
      \begin{column}{0.5\textwidth}
        \begin{align*}
          \synt{les} &: \textcolor{red}{N \limp NP} \\
          \synt{aime} &: \textcolor{red}{NP \limp NP \limp S}
        \end{align*}
      \end{column}
    \end{columns}

    $$\synt{aime}\ \synt{Laura}\ (\synt{les}\ \synt{tatous}) :
    \textcolor{darkgreen}{S}$$
  \end{block}

  \begin{block}{Object languages}
    A projection of an abstract language with a \textcolor{darkgold}{lexicon}.

    \begin{columns}[t]
      \begin{column}{0.5\textwidth}
        \begin{align*}
          \mathcal{L}(\synt{tatous}) &= \textcolor{darkgold}{tatous} \\
          \mathcal{L}(\synt{Laura}) &= \textcolor{darkgold}{Laura}
        \end{align*}
      \end{column}
      \begin{column}{0.5\textwidth}
        \begin{align*}
          \mathcal{L}(\synt{les}) &= \textcolor{darkgold}{\lambda^{\circ} x.\ les + x} \\
          \mathcal{L}(\synt{aime}) &= \textcolor{darkgold}{\lambda^{\circ} x y.\ x + aime + y}
        \end{align*}
      \end{column}
    \end{columns}

    $$\mathcal{L}(\synt{aime}\ \synt{Laura}\ (\synt{les}\ \synt{tatous}))
    =_{\beta} Laura + aime + les + tatous$$
  \end{block}
\end{frame}

\note[itemize]{\item Abstract Categorial Grammars introduce two families of
  languages: the so-called abstract languages and object languages.

\item Abstract languages are sets of lambda terms built from some signature of
  typed constants where every term of the language is well-typed and has a
  specific distinguished type. Here we have an example signature giving some
  familiar types to French words and a term of type S which corresponds to the
  structure of a French sentence.

\item Object languages are then derived from abstract languages by replacing
  all the constants in their terms with specific substitutes given by a
  lexicon. Here we have such a lexicon that when applied to the example
  sentence structure, it will yield a term that beta-reduces to the French
  sentence ``Laura aime les tatous''.

\item An ACG is then usually represented as a quadruple consisting of an
  abstract signature, a distinguished abstract type, an object signature and a
  lexicon connecting the two signatures.

\item 0:35}

\begin{frame}
  \frametitle{ACG Diagrams}

  \begin{center}
    \includegraphics[height=0.8\textheight]{../diagrams/double-acg.pdf}
  \end{center}
\end{frame}

\note[itemize]{\item Outside of language-theoretical explorations, considering
  single ACGs is seldom useful. In this typical example, we have two ACGs that
  share a single abstract language of sentence structures that is mapped via
  two different lexicons to phonological strings and semantic formulae.

\item Instead of talking about two quadruples that share some elements, we
  will prefer the more eloquent diagrams.

\item 0:20}

\begin{frame}
  \frametitle{ACG Patterns}

  \begin{columns}[c]
     \begin{column}{0.5\textwidth}
      \begin{block}{Constraint}
        \vspace{2 mm}
        \begin{center}
          \includegraphics[height=0.7\textheight]{../diagrams/serial-over-parallel-slides.pdf}
        \end{center}
      \end{block}
    \end{column}
    \begin{column}{0.5\textwidth}
      \begin{block}{Elaboration}
        \vspace{2 mm}
        \begin{center}
          \includegraphics[height=0.7\textheight]{../diagrams/parallel-over-serial-slides.pdf}
        \end{center}
      \end{block}
    \end{column}
 \end{columns}
\end{frame}

\note[itemize]{\item Besides the diagram we just saw, there are other
  established patterns in ACGs.

\item The Constraint pattern lets us retroactively restrain the members of
  previously defined languages. Whereas before, any sentence structure that
  had the distinguished type was considered valid, with the added constraint,
  we only consider those that can be generated from the newly added signature
  and lexicon. The new signature usually mimics the old one but refines its
  types so that certain terms will no longer be admissible.

\item The Elaboration pattern serves a different purpose. It lets us decompose
  an involved translation into two smaller translations and expose the
  intermediate product. In this case, we split the signature of sentence
  structures into a more superficial analytic layer which accounts only for
  syntactic ambiguities and a deeper tectogrammatical layer which goes
  further, into the semantic ambiguities of scope.

\item 0:45}

\section{Why Graphical ACGs?}

\note{Now that we are familiar with ACGs, I will proceed to show the problems
  that motivated our proposed extension.}

\begin{frame}
  \frametitle{Composing ACG Patterns I}

  \begin{columns}[c]
    \begin{column}{0.5\textwidth}
      \begin{center}
        \includegraphics[height=0.7\textheight]{../diagrams/fail1.pdf}
      \end{center}
    \end{column}
    \begin{column}{0.5\textwidth}
      \begin{center}
        \includegraphics[height=0.7\textheight]{../diagrams/fail2.pdf}
      \end{center}
    \end{column}
  \end{columns}
\end{frame}

\note[itemize]{{\scriptsize
\item Our problems revolve around the fact that all the diagrams drawn for
  ACGs are arborescences. There is always a single abstract language and all
  the other languages are established by projecting that abstract language
  with some composition of lexicons. This limits us when we try to combine the
  two patterns we just saw in a single grammar.

\item Consider the situation where we want to distinguish the analytic syntax
  and tectogrammatical levels of sentence structure, while adding a constraint
  on the analytic level. One way we can try to do this is to have the
  constraint signature constrain the analytic syntax level directly. But then
  the tectogrammatical level, which provides the link between syntax and
  semantics, has to deal with the constraint signature by having to generate
  correct members of its language. Thus we have an unwarranted interaction
  between internal details of syntax and the syntax-semantics interface.

}{\scriptsize
\item Another solution would be to keep the syntax-semantics interface pure by
  having it still connect just the analytic syntax with the
  semantics. However, then the constraint has to be implemented on top of all
  that, meaning that the constraint now has to deal with matters of
  tectogrammatics and the syntax-semantics interface. Again, we arrive at
  accidental complexity, this time with the syntax-semantics interface
  contaminating the syntactic constraint.

\item 0:54}}

\begin{frame}
  \frametitle{Composing ACG Patterns II}

  \begin{columns}[c]
    \begin{column}{0.5\textwidth}
      \begin{center}
        \includegraphics[height=0.7\textheight]{../diagrams/constr-fail-1.pdf}
      \end{center}
    \end{column}
    \begin{column}{0.5\textwidth}
      \begin{center}
        \includegraphics[height=0.7\textheight]{../diagrams/constr-fail-2.pdf}
      \end{center}
    \end{column}
  \end{columns}
\end{frame}

\note[itemize]{{\footnotesize \item When writing a wide-coverage grammar, we
    will be especially interested in implementing many different constraints
    and we would like to try to factor them out, maybe by using the Constraint
    pattern. However, this proves to be difficult thanks to the homomorphism
    property of lexicons, which lies at the heart of ACGs. The consequence of
    the homomorphism property that's interesting to us is that the image of a
    well-typed term with a lexicon must be another well-typed term.

\item This means that when we try to attach a second constraint in the example
  on the left, its types have to essentially ensure that all the licensed
  structures not only satisfy the new constraint, but also the existing one.

\item This problem is exacerbated by iteration. When adding a third
  constraint, we are obliged to engineer its signature such that the licensed
  terms conform to all three constraints. In the end, there is not much
  redeeming value to stacking constraints like this and we end up implementing
  multiple constraints in a single signature which, as we'll see, can get very
  complex.

\item 0:34}}

\begin{frame}
  \frametitle{Multiple Constraints}
  Agreement:
    \begin{align*}
      A_{aime_1} &: NP_{\textcolor{darkgold}{NUM{=}SG}} \limp NP_{\textcolor{darkgold}{NUM{=}SG}} \limp S \\
      A_{aime_2} &: NP_{\textcolor{darkgold}{NUM{=}SG}} \limp NP_{\textcolor{darkgold}{NUM{=}PL}} \limp S
    \end{align*}

  Extraction:
    \begin{align*}
      E_{aime_1} &: NP_{\textcolor{darkgreen}{VAR{=}F}} \limp NP_{\textcolor{darkgreen}{VAR{=}F}} \limp S_{\textcolor{darkgreen}{EXT{=}NO}} \\
      E_{aime_2} &: NP_{\textcolor{darkgreen}{VAR{=}T}} \limp NP_{\textcolor{darkgreen}{VAR{=}F}} \limp S_{\textcolor{darkgreen}{EXT{=}ROOT}} \\
      E_{aime_3} &: NP_{\textcolor{darkgreen}{VAR{=}F}} \limp NP_{\textcolor{darkgreen}{VAR{=}T}} \limp S_{\textcolor{darkgreen}{EXT{=}ROOT}
}    \end{align*}

  Combined:
  \begin{small}
  \begin{align*}
      C_{aime_1} &: NP_{\textcolor{darkgold}{NUM{=}SG},\textcolor{darkgreen}{VAR{=}F}} \limp NP_{\textcolor{darkgold}{NUM{=}SG},\textcolor{darkgreen}{VAR{=}F}} \limp S_{\textcolor{darkgreen}{EXT{=}NO}} \\
      C_{aime_2} &: NP_{\textcolor{darkgold}{NUM{=}SG},\textcolor{darkgreen}{VAR{=}T}} \limp NP_{\textcolor{darkgold}{NUM{=}SG},\textcolor{darkgreen}{VAR{=}F}} \limp S_{\textcolor{darkgreen}{EXT{=}ROOT}} \\
      C_{aime_3} &: NP_{\textcolor{darkgold}{NUM{=}SG},\textcolor{darkgreen}{VAR{=}F}} \limp NP_{\textcolor{darkgold}{NUM{=}SG},\textcolor{darkgreen}{VAR{=}T}} \limp S_{\textcolor{darkgreen}{EXT{=}ROOT}} \\
      C_{aime_4} &: NP_{\textcolor{darkgold}{NUM{=}SG},\textcolor{darkgreen}{VAR{=}F}} \limp NP_{\textcolor{darkgold}{NUM{=}PL},\textcolor{darkgreen}{VAR{=}F}} \limp S_{\textcolor{darkgreen}{EXT{=}NO}} \\
      C_{aime_5} &: NP_{\textcolor{darkgold}{NUM{=}SG},\textcolor{darkgreen}{VAR{=}T}} \limp NP_{\textcolor{darkgold}{NUM{=}PL},\textcolor{darkgreen}{VAR{=}F}} \limp S_{\textcolor{darkgreen}{EXT{=}ROOT}} \\
      C_{aime_6} &: NP_{\textcolor{darkgold}{NUM{=}SG},\textcolor{darkgreen}{VAR{=}F}} \limp NP_{\textcolor{darkgold}{NUM{=}PL},\textcolor{darkgreen}{VAR{=}T}} \limp S_{\textcolor{darkgreen}{EXT{=}ROOT}
}  \end{align*}
  \end{small}
\end{frame}

\note[itemize]{\item Here we have an example of trying to combine the
  treatment of two independent constraints in a single signature. In the first
  fragment, we see a few types refined to handle agreement in number. In the
  second, we have the same wordform in a refinement which enforces a
  constraint on extraction and relative clauses. Finally, in the third block,
  we have a slice of a signature which enforces both constraints at the same
  time.

\item While the individual constraint signatures are reasonably
  understandable, the combined signature grows exponentially with the number
  of features and it becomes difficult to see whether constraints are
  independent of each other or not.

\item 0:29}

\begin{frame}
  \frametitle{Graphical Abstract Categorial Grammars (G-ACGs)}

  \begin{columns}[c]
    \begin{column}{0.6\textwidth}
      \begin{center}
        \includegraphics[height=0.7\textheight]{../diagrams/final-slides.pdf}
      \end{center}
    \end{column}

    \begin{column}{0.4\textwidth}
        \begin{itemize}
        \item DAG
        \item Nodes:
          \begin{itemize}
          \item Signatures
          \item Distinguished types
          \end{itemize}
        \item Edges:
          \begin{itemize}
          \item Lexicons
          \end{itemize}
        \end{itemize}
    \end{column}
  \end{columns}
\end{frame}

\note[itemize]{\item The framework that we would like to work in would ideally
  let us build a grammar like this. Individual constraints are laid
  side-by-side without any cumulative complexity and outside of the scope of
  the syntax-semantics interface. With no directed paths leading between the
  constraint or tectogrammatics signatures, there are no direct dependencies
  between signatures which handle different issues in our grammar.

\item What we have here is the basic notation of our Graphical Abstract
  Categorial Grammars: a directed acyclic graph whose nodes are labelled with
  signatures and distinguished types and whose edges are labelled with
  lexicons. Having introduced the notation, we are now obliged to answer the
  question of how this notation establishes languages.

\item 0:36}

\section{Semantics of G-ACG}

\begin{frame}
  \frametitle{Two Intuitions for G-ACGs}

  \setbeamercolor{block title}{bg=limeblue}
  \setbeamercolor{block body}{bg=limeblue}
  \begin{block}{Algebraic Perspective}
    \begin{itemize}
    \item Nodes as \emph{languages}
    \end{itemize}
  \end{block}
  \setbeamercolor{block title}{bg=limegreen}
  \setbeamercolor{block body}{bg=limegreen}
  \begin{block}{Generative Perspective}
    \begin{itemize}
    \item Nodes as \emph{terms}
    \end{itemize}
  \end{block}
\end{frame}

\note[itemize]{\item We will present two interpretations of the G-ACG
  notation.

\item In the first, which we dubbed ``algebraic'', nodes are identified with
  languages and edges tell us how languages are defined in terms of each other
  using lexicons and intersections

\item In the second interpretation, the ``generative'' one, nodes are seen as
  standing for individual terms and edges are seen as statements that applying
  a particular lexicon to one term produces another. The languages are then
  established through the notion of a model.

\item 0:31}

\begin{frame}
  \frametitle{Algebraic Intuition}

  \begin{columns}[c]
    \begin{column}{0.7\textwidth}
      \setbeamercolor{block title}{bg=limeblue}
      \setbeamercolor{block body}{bg=limeblue}
      \begin{block}{}
        \begin{itemize}
        \item Node = Language
        \item Edge = Projection with a \textcolor{darkgreen}{lexicon}
        \item Two incoming edges = \textcolor{red}{Intersection} of languages
        \end{itemize}
      \end{block}

$$
\mathcal{I}_{\mathcal{G}}(v) = \{t \in \Lambda(\Sigma_v)
\mid\ \vdash_{\Sigma_v} t : S_v\}
$$
$$
\mathcal{E}_{\mathcal{G}}(v) = \mathcal{I}_{\mathcal{G}}(v) \cap
\textcolor{red}{\bigcap_{(u,v) \in E}}
\textcolor{darkgreen}{\mathcal{L}_{(u,v)}}(\mathcal{E}_{\mathcal{G}}(u)).
$$
    \end{column}
    \begin{column}{0.3\textwidth}
      \includegraphics[width=\textwidth]{../diagrams/algebraic-perspective.pdf}
    \end{column}
  \end{columns}

\end{frame}

\note[itemize]{\item In the algebraic paradigm, we associate a language with
  every node of the G-ACG. A single incoming edge tells us that the target
  language is the projection of the source language with the edge's
  lexicon. When multiple incoming edges meet, we take the intersection of the
  projections.

\item Formally, this is realized by establishing intrinsic languages for every
  node which correspond exactly to the abstract languages of ACGs; they are
  simply sets of terms having some distinguished type. This then helps us
  inductively build the extrinsic languages, which are defined as the
  intersection of the lexical projections of the node's predecessors'
  extrinsic languages, or the node's intrinsic language if it has no
  predecessors.

\item 0:43}

\begin{frame}
  \frametitle{Two Problems with the Algebraic View}

  \begin{columns}[t]
    \begin{column}{0.5\textwidth}
      \begin{center}
        \includegraphics[height=0.6\textheight]{../diagrams/both-patterns-slides.pdf}
      \end{center}

      How do we get the language of meanings expressible only by
      syntactically correct sentences?
    \end{column}
    \begin{column}{0.5\textwidth}
      \begin{center}
        \includegraphics[height=0.6\textheight]{../diagrams/diamond-grammar.pdf}
      \end{center}

      How do we express the language of $Bottom$ terms whose $Left$ and
      $Right$ antecedents are generated by the same $Top$ antecedent?
    \end{column}
  \end{columns}
\end{frame}

\note[itemize]{{\scriptsize \item The algebraic interpretation that we
    developed can be counter-intuitive in certain situations. First, consider
    the combination of a syntactic constraint that is independent of the
    syntax-semantics interface, as can be seen in the diagram on the
    left. When we take a semantic formula belonging to the extrinsic language
    of the node $Sem$, we can take its antecedent in the syntax-semantics
    interface $SyntSem$ and project it to find the syntactic counterparts of
    our semantic formula. However, we have no guarantee that these
    counterparts will satisfy the syntactic constraint, in other words, they
    might lack an antecedent in the $Constr$ signature. Thus, it becomes
    difficult to establish the language of meanings which are expressible only
    by syntactically correct sentences since the extrinsic language of $Sem$
    is not constrained by the $Constr$ signature and thus overgenerates in
    this case.

\item Another counter-intuitive example is the diamond grammar on the
  right. If we take a member of the $Bottom$ extrinsic language, it must have,
  by definition, antecedents in the $Left$ and $Right$ extrinsic
  languages. These two antecedents then must have their own antecedents in the
  $Top$ extrinsic language. However, they don't have to be the same
  antecedent as the single shared node on the diagram might suggest.

\item 0:58}}

\begin{frame}
  \frametitle{Generative Intuition}

  \begin{columns}[c]
    \begin{column}{0.7\textwidth}
\setbeamercolor{block title}{bg=limegreen}
      \setbeamercolor{block body}{bg=limegreen}
      \begin{block}{}
        \begin{itemize}
        \item Node = \textcolor{red}{Term}
        \item Edge = Image with a \textcolor{darkgreen}{lexicon}
        \item Two incoming edges = \textcolor{red}{Equality} of images
        \end{itemize}
      \end{block}

      \vspace{3 mm}

      $t \in \mathcal{P}_{\mathcal{G}}(u)$ iff $\exists T :
      \textcolor{red}{V(G) \to \Lambda}$ s.t.:
      \begin{itemize}
      \item $T_u = t$.
      \item $\forall v \in V(G)$, $\vdash_{\Sigma_v} T_v : S_v$.
      \item $\forall (v,w) \in E(G)$,
        $\textcolor{darkgreen}{\mathcal{L}_{(v,w)}(T_v) = T_w}$.
      \end{itemize}
    \end{column}
    \begin{column}{0.3\textwidth}
      \includegraphics[width=\textwidth]{../diagrams/generative-perspective.pdf}
    \end{column}
  \end{columns}
\end{frame}

\note[itemize]{\item We now turn to another interpretation of G-ACGs, one that
  we called the ``generative'' interpretation, though ``logical'' or
  ``relational'' are also apt names.

\item The nodes of the graph stand for terms, that is, they are term variables
  that range over the nodes' intrinsic languages. An edge is a statement
  saying that the pair of terms must belong to the edge's lexicon, which is to
  say that applying the lexicon to the source term must yield the target term.

\item We then consider a model of this graphical formula to be a labelling
  which assigns terms to the vertices such that both the domain restrictions
  and the edge statements are satisfied. From this we establish the
  pangraphical language of a node as the set of all the possible terms that
  can label the node in a valid model.

\item 0:44}

\begin{frame}
  \frametitle{Algebraic and Generative Relations}

  \note{We have presented two interpretations of G-ACGs based on two different
    intuitive readings of ACG diagrams, one where nodes are thought of as
    languages and one where they stand for individual terms. The result shown
    here validates these two ways of reasoning about diagrams in the context
    of classical ACGs since they lead to the same languages as the canonical
    definition of an object language.

    0:25}

  \begin{itemize}
  \item $\forall \, \mathcal{G}, u$ where $\mathcal{G}$ is arborescent:
    $$\mathcal{E}_{\mathcal{G}}(u) = \mathcal{P}_{\mathcal{G}}(u)
    = \mathcal{O}_{\mathcal{G}}(u)$$
  \pause
  \item However, $\exists \, \mathcal{G}, u$ such that:
    $$\mathcal{E}_{\mathcal{G}}(u) \neq \mathcal{P}_{\mathcal{G}}(u)$$
  \end{itemize}
\end{frame}

\note{However, as soon as we turn to G-ACGs which are not arborescent, the
  difference between these two perspectives becomes substantial since the two
  definitions we elaborated on before lead to different languages.

  0:14}

\begin{frame}
  \frametitle{Properties of G-ACG Languages}

  $$\forall \, \mathcal{G}, u. \; \mathcal{I}_{\mathcal{G}}(u) \supseteq
  \mathcal{E}_{\mathcal{G}}(u) \supseteq \mathcal{P}_{\mathcal{G}}(u)$$

  $$\mathcal{I} \subseteq \mathcal{E} \subseteq \mathcal{P}$$

  $$\mathcal{I} = \mathcal{A}$$

  $$\mathcal{O} = \mathcal{A}^{\mathcal{L}}$$

  $$\mathcal{E} = \mathcal{A}^{\mathcal{L}{\cap}}$$
\end{frame}

\note[itemize]{{\scriptsize \item We now proceed to give some basic properties
    of the language families we have just introduced.

\item The three language families that we have just defined (intrinsic,
  extrinsic and pangraphical) were increasingly restrictive. The intrinsic
  languages were constrained only by the node's signature and distinguished
  type, whereas the extrinsic languages were constrained also by the node's
  predecessors and the pangraphical languages were constrained by the entire
  graph.

\item We can prove that any extrinsic language is also pangraphical and
  therefore we get that the three language families we introduced were also
  increasingly expressive.

\item Intrinsic languages were useful to establish extrinsic and pangraphical
  languages. They are exactly the same thing as the abstract languages of
  ACGs.

\item Extrinsic languages, even though they can give counter-intuitive
  interpretations in some cases, are still very interesting. It can be shown
  that they extend the expressive power of object languages precisely by
  introducing closure on intersection and nothing more.

}{\tiny
\item We don't have a similar result for pangraphical languages but we find
  them useful since they more closely capture our intuitions about ACG
  diagrams and so, for example, they resolve the two problems we saw with
  interpreting G-ACGs as extrinsic languages.

\item 1:01}}

\section{Illustration}

\note{We now illustrate the kind of grammar for which we introduced this
  extension.}

\begin{frame}
  \frametitle{Basic Setup}
  \begin{center}
    \includegraphics[height=0.8\textheight]{../diagrams/illustration0.pdf}
  \end{center}
\end{frame}

\note{We start with the canonical example of a tectogrammatical language that
  is projected into phenogrammatical strings and semantic formulae.

  0:09}

\begin{frame}
  \addtocounter{framenumber}{-1}
  \frametitle{Single Constraint}
  \begin{center}
    \includegraphics[height=0.8\textheight]{../diagrams/illustration1.pdf}
  \end{center}
\end{frame}

\note{We then add one of the three constraints on extraction developed by
  Sylvain Pogodalla and Florent Pompigne in their 2010 paper.

  0:08}

\begin{frame}
  \addtocounter{framenumber}{-1}
  \frametitle{Multiple Constraints}
  \begin{center}
    \includegraphics[height=0.8\textheight]{../diagrams/illustration2.pdf}
  \end{center}
\end{frame}

\note{Then we add the other two constraints the same way. It is here that we
  leave the scope of what classical ACGs let us do.

  0:09}

\begin{frame}
  \addtocounter{framenumber}{-1}
  \frametitle{Exposing Another Signature}
  \begin{center}
    \includegraphics[height=0.8\textheight]{../diagrams/illustration3.pdf}
  \end{center}
\end{frame}

\note{Next, we proceed by elaborating the syntactic lexicon, again by
  literally copying existing work from one of Sylvain Pogodalla's papers. This
  exposes a new signature of lower-order syntactic structures which are more
  comfortable to work with.

  0:14}

\begin{frame}
  \addtocounter{framenumber}{-1}
  \frametitle{Constraints on Multiple Signatures}
  \begin{center}
    \includegraphics[height=0.8\textheight]{../diagrams/illustration4.pdf}
  \end{center}
\end{frame}

\note{We then exploit this new syntactic signature by adding constraints. We
  refine some types to simulate Guy Perrier's treatment of negative
  determiners in French. We also add a constraint for enforcing agreement of
  number.

  0:13}

\begin{frame}
  \frametitle{Reflection}
  \begin{columns}[c]
    \begin{column}{0.5\textwidth}
      \begin{center}
        \includegraphics[height=0.8\textheight]{../diagrams/illustration5.pdf}
      \end{center}
    \end{column}
    \begin{column}{0.5\textwidth}
      \begin{itemize}
        \item 9 signatures
        \item 8 lexicons
        \item High level of redundancy
        \item Constraints decomplected from each other
        \item Constraints on different signatures
      \end{itemize}
    \end{column}
  \end{columns}
\end{frame}

\note{One thing to remark about this grammar is its size. Having a different
  signature for each constraint means that there is quite a lot of redundancy
  since we are repeatedly refining the same signature in different ways. On
  the other hand, we have segregated all of our constraints into individual
  signatures, keeping them simple. Furthermore, the constraints are attached
  to different nodes of the G-ACG, making the introduction of new constraints
  flexible in how deeply they want to get involved with the rest of the
  grammar.

  0:21}

\section{Conclusion}

\begin{frame}
  \frametitle{Recapitulation}

  \begin{itemize}
  \item ACG diagrams as mathematical objects
  \vfill
  \item Non-arborescence enables intersection
  \vfill
  \item Intersection enables conjunction of independent constraints
  \vfill
  \item Large-scale grammar: collection of fragments
  \end{itemize}
\end{frame}

\note[itemize]{\item We started by reinterpreting ACGs by treating their
  diagrams as primary defining constructs instead of as artifacts used only
  for illustration.

\item This let us explore the semantics of non-arborescent diagrams
  directly. Through our explorations, we have arrived at a formalism which
  facilitates intersection.

\item Intersection turned out to be a key feature for the arbitrary
  conjunction of independent constraints, which is useful in crafting detailed
  grammars.

\item We illustrated our approach by building a grammar which enforces a small
  but diverse range of constraints and is broken up into manageable
  chunks. This modularity was showcased by assembling the grammar in large
  part by simply copying existing work without any modification.

\item 0:35}

\begin{frame}
  \frametitle{Future Directions}

  \begin{itemize}
  \item Are ACG object languages closed on
    intersection? $$\mathcal{O}^{\mathcal{L}{\cap}} \subseteq \mathcal{O}^{\mathcal{L}}$$
  \item Are extrinsic languages as expressive as pangraphical
    ones? $$\mathcal{P} \subseteq \mathcal{E}$$
  \item Is there a metagrammatical approach that makes refining the types of a
    signature more concise?
  \vfill
  \item Is this enough to build a satisfactory grammar of a real language?
  \end{itemize}
\end{frame}

\note[itemize]{{\scriptsize \item Are ACG object languages closed on
    intersection? If this were true, then our proposal would be just a
    conservative extension which makes conjunction of constraints easier. If
    this were false, then our proposal is necessary to make the languages
    definable by the formalism closed on intersection, which we see as
    critical for conjoining independent constraints.

\item We like that we can exactly describe the set of extrinsic languages
  w.r.t. ACG object languages. However, we still prefer pangraphical languages
  as we find them more intuitive and easier to understand. We already know
  that extrinsic languages are pangraphical. Proving the converse would be one
  way of characterizing the set of pangraphical languages.

\item A constraint signature that is created by refining the types shares a
  lot of similarities with the original signature. Likewise, constraint
  signatures developed using this techniques are alike. This makes us think
  that it should be possible to find some kind of meta-notation that will be
  able to convey new constraints more concisely than refining an entire
  signature by hand.

\item Finally, we are also curious to see how far we can get in building G-ACG
  grammars before we run into another problem that will warrant extending the
  formalism in yet another direction.

\item 1:07}}

\end{document}