### 2013-12-25 sjtu/body/ack.tex

%%==================================================
%% thanks.tex for SJTU Master Thesis
%% based on CASthesis
%% modified by wei.jianwen@gmail.com
%% version: 0.3a
%% Encoding: UTF-8
%% last update: Dec 5th, 2010
%%==================================================

\begin{thanks}

  I thank the European Union and the Erasmus Mundus Language and Communication
  Technologies Master program for funding my studies.

  I thank my supervisors, Maxime Amblard in Nancy and Zhao Hai in Shanghai,
  for their leadership and assistance.

  I thank my colleague Sai Qian for helping me with the Chinese translations.

  I thank my friends in Shanghai and Nancy for two wonderful years.

  I thank my parents in supporting me in everything I chose to do.

  Thank you!

\end{thanks}