### 2014-02-18 lenls/post-proc.tex
POST: Final remarks from Maxime, flipped sigs in story diagrams

\documentclass{llncs}
%
\usepackage{makeidx}

\usepackage{caption}
\usepackage[labelformat=simple]{subcaption}
\renewcommand\thesubfigure{(\alph{subfigure})}

\usepackage[utf8]{inputenc}
\usepackage{amsmath}
\usepackage{xcolor}
\usepackage{graphicx}
\usepackage{gb4e}
\usepackage{bussproofs}
\usepackage{hyperref}
\usepackage{cancel}
\usepackage{pdfpages}
\usepackage{enumitem}

%% Linear implication from Alessio Guglielmi
%% https://groups.google.com/forum/?fromgroups=#!topic/comp.text.tex/0B4C3F_BsVI
\def\limp {\mathbin{{-}\mkern-3.5mu{\circ}}}

%
\begin{document}
%
\frontmatter          % for the preliminaries
%
\pagestyle{headings}  % switches on printing of running heads

\mainmatter              % start of the contributions
%
\title{Integration of Multiple Constraints in ACG}
%
\titlerunning{Constraints in ACG}  % abbreviated title (for running head)
%                                    also used for the TOC unless
%                                    \toctitle is used
%
\author{Jiří Maršík\inst{1,2,3} \and Maxime Amblard\inst{1,2,3}}
%
\authorrunning{Maršík et al.} % abbreviated author list (for running head)
%
%%%% list of authors for the TOC (use if author list has to be modified)
\tocauthor{Jiří Maršík and Maxime Amblard}
%
\institute{Université de Lorraine, Laboratoire lorrain de recherche en informatique et ses applications, UMR 7503, Vandoeuvre-lès-Nancy, 54500, France\\
\and
INRIA, Villers-lès-Nancy, 54600, France\\
\and
CNRS, Loria, Vandoeuvre-lès-Nancy, UMR 7503, 54500, France\\
\email{\{jiri.marsik, maxime.amblard\}@loria.fr}}

\maketitle              % typeset the title of the contribution

\begin{abstract}
This proposal is a first step towards a wide-coverage Abstract Categorial
Grammar (ACG) that could be used to automatically build discourse-level
representations. We focus on the challenge of integrating the treatment of
disparate linguistic constraints in a single ACG and propose a generalization
of the formalism: Graphical Abstract Categorial Grammars.
\keywords{abstract categorial grammars, grammar engineering, grammatical formalisms, formal grammars, computational linguistics}
\end{abstract}

Abstract Categorial Grammars (ACGs) \cite{de2001towards} have shown to be a
viable formalism for elegantly encoding the dynamic nature of
discourse. Proposals based on continuation semantics \cite{de2006towards} have
tackled topics such as event anaphora \cite{qian2011event}, SDRT discourse
structure \cite{asher2011sdrt} and modal accessibility constraints
\cite{asher2011montagovian}. However, all of these treatments only consider
tiny fragments of languages. We are interested in building a wide-coverage
grammar which integrates and reconciles the existing formal treatments of
discourse and allows us to study their interactions and to build discourse
representations automatically.

A presupposed condition to actually using an ACG to describe discourse in a
large scope is to have a large scale ACG grammar in the first place. The work
presented here was motivated by this quest for a wide-coverage ACG grammar, as
seen from the point of view of language as a system of largely orthogonal
constraints. Encoding ancillary constraints in a type system can make the
system hard to read and understand. If multiple constraints written in the
style of \cite{pogodalla2012controlling} are to be enforced in the same
grammar, we advocate for extending the formalism to prevent the incidental
complexity that would otherwise emerge.

We begin with a short review of Abstract Categorial Grammars where we
highlight the ways in which Abstract Categorial Grammars facilitate
decomposition of linguistic description. We follow that by a closer look at an
example of a linguistic constraint. This discussion motivates our proposed
extension. We then present the Graphical Abstract Categorial Grammars and
consider some of its formal properties. We then finish with an illustration of
our approach which incorporates linguistic contributions from multiple sources
into a single coherent and simple grammar and examine the limitations and
challenges of adopting this approach.

\section{Abstract Categorial Grammars}
\label{sec:acg-comp}

We first review the grammatical framework in which we conduct our
work. Abstract Categorial Grammars are built upon two mathematical structures:
\emph{(higher-order) signatures} and \emph{lexicons}.

\subsection{Higher-Order Signatures}
\label{ssec:sig}

A \textbf{higher-order signature} is a set of elements that we call
\emph{constants}, each of which is associated with a type. Formally, it
is defined as a triple $\Sigma = \mathopen{<}A, C, \tau\mathclose{>}$,
where:
\begin{itemize}
  \item $C$ is the (finite) set of constants
  \item $A$ is a (finite) set of atomic types
  \item $\tau$ is the type-associating mapping from $C$ to
    $\mathcal{T}(A)$, the set of types built over $A$
\end{itemize}

In our case, $\mathcal{T}(A)$ is the implicative fragment of linear and
intuitionistic logic with $A$ being the atomic propositions. This means
that $\mathcal{T}(A)$ contains all the $a \in A$ and all the $\alpha \limp
\beta$ and $\alpha \to \beta$ for $\alpha, \beta \in \mathcal{T}(A)$.

A signature $\Sigma = \mathopen{<}A, C, \tau\mathclose{>}$, by itself, already
lets us define an interesting set of structures, that is the set
$\Lambda(\Sigma)$ of \emph{well-typed lambda terms} built upon the signature
$\Sigma$. To make this structure even more useful, we often focus ourselves
only on terms that have a specific \emph{distinguished type}. Using this
notion of a signature of typed constants and some distinguished type, we can
describe languages of, e.g. tree-like (and by extension string-like), lambda
terms.

\subsection{Lexicons}

The idea of a signature is coupled with the one of \textbf{lexicon},
which is a mapping between two different signatures (mapping the
constants of one into well-typed terms of the other). Formally speaking,
a lexicon $\mathcal{L}$ from a signature $\Sigma_1 = \mathopen{<}A_1,
C_1, \tau_1\mathclose{>}$ (which we call the abstract signature) to a
signature $\Sigma_2 = \mathopen{<}A_2, C_2, \tau_2\mathclose{>}$ (which
we call the object signature) is a pair $\mathopen{<}F, G\mathclose{>}$
such that:

\begin{itemize}
\item $G$ is a mapping from $C_1$ to $\Lambda(\Sigma_2)$ assigning to
  every constant of the abstract signature a term in the object
  signature, which can be understood as its
  interpretation/implementation/realization.
\item $F$ is a mapping from $A_1$ to $\mathcal{T}(A_2)$ which links the
  abstract-level types with the object-level types that they are
  realized as.
\item $F$ and $G$ are compatible, meaning that for any $c \in C_1$, we
  have $\vdash_{\Sigma_2} G(c) : \hat{F}(\tau_1(c))$ (we will be using
  $\hat{F}$ and $\hat{G}$ to refer to the homomorphic extensions of $F$
  and $G$ to $\mathcal{T}(A_1)$ and $\Lambda(\Sigma_1)$
  respectively).
\end{itemize}

Now we have enough machinery in hand to describe how ACGs define
languages. Given some signature and a distinguished type, we can talk of the
\emph{abstract language}, which is the set of well-typed terms built on the
signature and having the distinguished type. If we then also consider a
lexicon and its abstract signature, then the language we get by mapping the
abstract language of the signature through the lexicon is called an
\emph{object language}.

An ACG for us will then be just a collection of signatures with their
distinguished types and the lexicons connecting these signatures, which we
will write down in diagrams such as the ones in Figure~\ref{fig:acg-comp}.

The most common pattern will have us using two object signatures,
$\Sigma_{String}$ and $\Sigma_{Sem}$, for the surface forms of utterances
(strings) and their logical forms (logical propositions) and an abstract
signature, $\Sigma_{Synt}$, which is connected to both of the object
signatures via lexicons (as we can see in
Figure~\ref{fig:acg-comp-basic}). Parsing is then just a matter of inverting
the surface lexicon to get the abstract term and then applying to it the
logical lexicon. Generation is symmetric, we simply invert the logical lexicon
and apply the surface lexicon.

\subsection{Systems of ACGs}

\begin{figure}[t]
  \centering
  \begin{subfigure}[b]{0.25\textwidth}
    \centering
    \includegraphics[height=0.2\textheight]{../diagrams/double-acg.pdf}
    \caption{\label{fig:acg-comp-basic} Connecting form with meaning.}
  \end{subfigure}
  \qquad
  \begin{subfigure}[b]{0.25\textwidth}
    \centering
    \includegraphics[height=0.2\textheight]{../diagrams/serial-over-parallel.pdf}
    \caption{\label{fig:acg-comp-constr} Adding a constraint (Constraint).}
  \end{subfigure}
  \qquad
  \begin{subfigure}[b]{0.25\textwidth}
    \centering
    \includegraphics[height=0.2\textheight]{../diagrams/parallel-over-serial.pdf}
    \caption{\label{fig:acg-comp-sem} Distinguishing syntactic and
      semantic ambiguities (Elaboration).}
  \end{subfigure}
  \caption{\label{fig:acg-comp} Diagrams of systems of ACGs.}
\end{figure}

Besides the pattern in Figure~\ref{fig:acg-comp-basic}, there have been other
techniques of decomposing linguistic description in the ACG literature. Here
we will focus on two that are relevant to our work.

The first constitutes the Constraint pattern used throughout
\cite{pogodalla2012controlling}. Imagine we want our grammar to enforce a
particular linguistic constraint, say that tensed clauses form islands for
quantifier scope, and suppose we want to proceed by refining our types so as
to make linguistic structures violating the constraint untypable. We could
dive into our grammar and rewrite it to use the new types. However, we can
also keep the existing grammar intact and complement it with a signature and a
lexicon that translates that signature to the abstract-most signature of our
existing grammar (Figure~\ref{fig:acg-comp-constr}). Usually, the new
signature will resemble the prior abstract-most one, albeit with more
fine-grained types necessary to express the constraint in question.

This pattern is practical for examining linguistic constraints one-by-one and
demonstrating that the formalism is capable of expressing them
individually. However, this pattern does not serve us much in the context of
ACGs when we try to build a grammar which incorporates many of these at the
same time. This is due to the fact that after grafting on the first
constraint, its added signature now becomes the abstract-most signature in the
grammar. Any other constraint that would be added using the Constraint pattern
would have to be translated into this new signature. This would mean that the
type systems of newly added constraints would have to re-enforce all of the
constraints introduced before, therefore nullifying almost any benefit of the
pattern.

Another pattern we will briefly mention is the Elaboration pattern in which an
edge in the diagram (a lexicon) is expanded into two edges and an intermediate
node (signature). We can illustrate this pattern on Tree Adjoining Grammars
\cite{degroote02}. Suppose we have a signature describing TAG derivation trees
and a lexicon which maps them to their yield. We might want to make this
grammar richer, and arguably more synoptic, by elaborating the translation
from derivation trees to yields by introducing derived trees and describe
separately how both derivation trees produce derived trees and how derived
trees produce the yields. This way, we can make emerge a potentially useful
linguistic structure.

This pattern was used in \cite{pogodalla2007generalizing} to elaborate the
translation from the common Montague-style \cite{montague1973proper}
higher-order signature to its yield by introducing lower-order syntactic
structures more reminiscent of TAG or CFG trees. This helps make a clear
distinction between ambiguities caused by syntax and by semantics by
segregating the purely syntactic description in the $Syntax$ signature from
the issues of operator scope that have to be solved in the $SyntSem$ signature
(Figure~\ref{fig:acg-comp-sem}). However, in classical ACGs, we cannot exploit
this pattern as much as we would like. If we were to use this grammar as a
kernel on top of which we would like to express a syntactic constraint, we
would not be able to attach the constraint's signature directly to the
$Syntax$ node but we would instead need to attach it to
$SyntSem$\footnote{This is due to the fact that ACG diagrams are always
  arborescences, in which the root node represents the abstract language from
  whose terms are generated the terms of all the object languages.}. This
means that the syntactic constraint would end up being expressed in terms of
the types of the syntax-semantics interface ($SyntSem$) rather than just pure
syntax ($Syntax$), which only adds unnecessary complexity.

\section{The Problem of Multiple Constraints}
\label{sec:constraints}

We will consider several linguistic constraints that have been given
formal treatments in grammatical formalisms.

In French, negation is signalled by prepending the particle \emph{ne} to the
negated verb in conjunction with using a properly placed accompanying word,
such as a negative determiner, in one of the verb's arguments. This phenomenon
has been elegantly formalized in the Frigram interaction grammar
\cite{perrier2007french}.

\begin{exe}
  \ex \label{ex:aucun-shallow} Jean \textbf{ne} parle à \textbf{aucun} collègue. \\
      (Jean speaks to no colleague.)
  \ex \label{ex:aucun-deep-obj} Jean \textbf{ne} parle à la femme d'\textbf{aucun} collègue. \\
      (Jean speaks to the wife of no colleague.)
  \ex \label{ex:aucun-deep-subj} \textbf{Aucun} collègue de Jean \textbf{ne} parle à sa femme. \\
      (No colleague of Jean's speaks to his wife.)
\end{exe}

We see here that the negative determiner \emph{aucun} can be present in
the subject or the object of the negated verb and it can modify the
argument directly or be attached to one of its complements. Furthermore,
we note that omitting either the word \emph{ne} or the word \emph{aucun}
while keeping the other produces a sentence which is considered
ungrammatical.

This difference in syntactic behavior between noun phrases that contain a
negative determiner and those that do not has implications for our
grammar. Since two terms that have an identical type in an ACG signature can
be freely interchanged in any sentence, we are forced to assign two different
types to these two different kinds of noun phrases.

% NOTE: We could use a single constant for the two paired words.
% NOTE: We could use a higher-order type for the negative determiner
% which demands the negative particle itself.

This leads us to a grammar in which we subdivide the atomic types $N$
and $NP$ into subtypes that reflect whether or not they contain a
negative determiner inside. Types of the other constants, such as the
preposition \emph{de} seen in (\ref{ex:aucun-deep-obj}) and
(\ref{ex:aucun-deep-subj}), will constrain their arguments to have
compatible features on their types and will propagate the information
carried in the features to its result type, e.g.:

\vspace{-4ex}

\begin{align*}
N_{de_1} &: NP_{NEG{=}F} \limp N_{NEG{=}F} \limp N_{NEG{=}F} \\
N_{de_2} &: NP_{NEG{=}F} \limp N_{NEG{=}T} \limp N_{NEG{=}T} \\
N_{de_3} &: NP_{NEG{=}T} \limp N_{NEG{=}F} \limp N_{NEG{=}T}
\end{align*}

\vspace{-1ex}

In the above, we elaborate the types $NP$ and $N$ with features ($NEG{=}F$
and $NEG{=}T$) and we give the different types for the preposition \emph{de}
in the fragment for negation ($N$). The types accept an $NP$ and an $N$ as
arguments with any combination of values for the feature $NEG$, except for the
case when both the prepositional $NP$ and the $N$ being modified both contain
free negative determiners (i.e. there is no $N_{de_4} : NP_{NEG{=}T} \limp
N_{NEG{=}T} \limp N_{NEG{=}?}$). This encodes a constraint that there can
be only one free negative determiner per phrase (free as in not hidden inside
an embedded clause). Besides this constraint, the types also dictate how the
information about negative determiners should propagate from the argument
constituent to the complex constituent (in this case, it is simple
disjunction).

Enforcing other constraints leads us to subdividing our ``atomic'' types even
further (e.g. the authors of \cite{pogodalla2012controlling} add features to
the $S$ and $NP$ types to implement constraints about extraction). Other
phenomena, such as agreement on morphological properties like number, gender,
person or case, intuitively lead us to make our types even more specific.

If we were to use this approach to write a grammar that enforces multiple
constraints at the same time, we would end up with complicated types, like the
one below, which provide complete specifications of the various possible
situations (in this grammar ($C$), the preposition \emph{de} has 12 different
types).

\vspace{-4ex}

$$
C_{de_{11}} :\ NP_{\textcolor{red}{NEG{=}T}, \textcolor{green}{VAR{=}F}, \textcolor{blue}{NUM{=}PL}}
\limp N_{\textcolor{red}{NEG{=}F},\textcolor{blue}{NUM{=}SG}}
\limp N_{\textcolor{red}{NEG{=}T},\textcolor{blue}{NUM{=}SG}}
$$

\vspace{-1ex}

This creates two problems. Firstly, the number of such types grows
exponentially with the number of features added. This can be fixed by
introducing dependent types into the type system as in
\cite{de2007type}. However, while this allows us to abstract over the
combinations of feature values and write our grammar down concisely, it does
not take away the complexity. The treatments of the various linguistic
phenomena are all expressed in the same types making it hard to see whether
they are independent or not. Since the treatments cannot be considered in
isolation, reasoning about the entire grammar becomes difficult and so does
enhancing it with more constraints. This is a fatal problem for a grammar
which strives to cover a wide range of linguistic facts. We firmly believe
that simplicity is a fundamental requirement for constructing a large and
robust grammar and our proposal aims to reclaim that simplicity.

In our grammar, we would like to combine several constraints
(Figure~\ref{fig:acg-comp-constr}), and possibly to also separate the
syntactic ambiguities from the purely semantic ones
(Figure~\ref{fig:acg-comp-sem}).  However, trying to mix these strategies in
the ACG framework forces us to solve all the constraints in a single type
signature or contaminate the syntax-semantics interface with the
implementation details of the syntactic layer, both of which introduce
incidental complexity we want to avoid.

We would like to have a system which would be characterized by a diagram like
the one on Figure~\ref{fig:gacg} where the constraint signatures delimit the
legal syntactic structures independently of each other and without interfering
with the syntax-semantics interface. However, ACG diagrams are limited to
arborescences and we are obliged to generalize them in order to get
the expected interpretation.

\section{Graphical ACGs}

\begin{figure}
  \centering
  \includegraphics[height=0.3\textheight]{../diagrams/final.pdf}
  \caption{\label{fig:gacg} A graphical ACG that implements two
      independent syntactic constraints and distinguishes syntactic and
      semantic ambiguities.}
\end{figure}

We define a \emph{graphical abstract categorial grammar} as a directed acyclic
graph (DAG) whose nodes are labeled with signatures (and distinguished types)
and whose edges are labeled with lexicons, in other words, a mathematical
reification of an ACG diagram that has been generalized from arborescences to
DAGs. We then search for an appropriate semantics for these structures,
i.e. how to determine what languages are defined by a graphical ACG.

\subsection{Nodes as Languages}

We first follow a paradigm in which nodes of the diagrams are interpreted as
languages with the edges telling us how these languages are defined in terms
of each other. A single arrow leading to a language means that the target
language is produced from the source by mapping it through a lexicon. We now
argue that the suitable meaning of two or more edges arriving at the same node
is intersection of languages based both on the simplicity of the resulting
definitions and on our expectations about the desired semantics.

In an ACG diagram, a node with no inbound edges stands for an abstract
language. This language is defined as the set of terms having the correct
type. If a node has an inbound edge, and therefore a parent, then the elements
of its language are also obliged to have an antecedent in the parent
language. It is a small step to go from this definition to the following: the
language of a node is the set of terms having the correct type and an
antecedent in the language of any of its parent nodes. This correctly
characterizes the current use of ACG diagrams, recognizing abstract languages
as a special case of object languages. Furthermore, this restatement
generalizes to DAGs and gives us the desired semantics for implementing
multiple constraints: intersection.

We can formalize the above definitions by introducing the notions of
\emph{intrinsic} and \emph{extrinsic} languages associated with some node $v$
in a graphical ACG $\mathcal{G}$:

$$
\mathcal{I}_{\mathcal{G}}(v) = \{t \in \Lambda(\Sigma_v)
\mid\ \vdash_{\Sigma_v} t : S_v\}
$$
$$
\mathcal{E}_{\mathcal{G}}(v) = \mathcal{I}_{\mathcal{G}}(v) \cap
\bigcap_{(u,v) \in E} \mathcal{L}_{(u,v)}(\mathcal{E}_{\mathcal{G}}(u))
$$

The intrinsic language is just the set of terms built on the node's signature
and having the node's distinguished type. The extrinsic language is
established by taking the extrinsic languages of its predecessors, mapping
them through the lexicons and taking their intersection, or just taking the
node's intrinsic language if it has no predecessors.

We then examine the relationship between the languages defined by ACGs
and graphical ACGs (G-ACGs). Intrinsic languages correspond exactly to
abstract languages and therefore the sets of languages definable by both
are equal.

$$
\mathcal{I} = \mathcal{A}
$$

G-ACG extrinsic languages correspond to ACG object languages with
intersection. More formally, while ACG object languages are ACG abstract
languages closed on transformation by a lexicon, G-ACG extrinsic languages are
ACG abstract languages closed on transformation by a lexicon and
intersection.

$$
\mathcal{O} = \mathcal{A}^{\mathcal{L}}
$$
$$
\mathcal{E} = \mathcal{A}^{\mathcal{L}{\cap}}
$$

\begin{proof}
  We prove $\mathcal{E} = \mathcal{A}^{\mathcal{L}{\cap}}$, or equivalently
  $\mathcal{E} = \mathcal{I}^{\mathcal{L}{\cap}}$, by proving $\mathcal{E}
  \subseteq \mathcal{I}^{\mathcal{L}{\cap}}$ and
  $\mathcal{I}^{\mathcal{L}{\cap}} \subseteq \mathcal{E}$.

  We begin with $\mathcal{E} \subseteq \mathcal{I}^{\mathcal{L}{\cap}}$. Let
  us consider any graphical ACG $\mathcal{G}$ and some topological ordering
  $v_1, \ldots, v_n$ of its nodes $V(G)$ (an ordering such that more abstract
  nodes always precede less abstract ones). We proceed by induction on the
  topological ordering and prove that $\mathcal{E}_{\mathcal{G}}(v_i) \in
  \mathcal{I}^{\mathcal{L}{\cap}}$ for all $v_i$ in $V(G)$:

  \begin{itemize}
  \item For $v_1$, the case is trivial. $v_1$ has no predecessors and so
    $\mathcal{E}_{\mathcal{G}}(v_1) = \mathcal{I}_{\mathcal{G}}(v_1) \in
    \mathcal{I}$.
  \item For any other node $v_n$, we inspect the definition of
    $\mathcal{E}_{\mathcal{G}}(v_n)$.

    $$\mathcal{E}_{\mathcal{G}}(v_n) = \mathcal{I}_{\mathcal{G}}(v_n) \cap
    \bigcap_{(u,v_n) \in E(G)}
    \mathcal{L}_{(u,v_n)}(\mathcal{E}_{\mathcal{G}}(u))$$

    By induction hypothesis, $\mathcal{E}_{\mathcal{G}}(u) \in
    \mathcal{I}^{\mathcal{L}{\cap}}$ for all the predecessors $u$ of
    $v_n$. Since, $\mathcal{I}^{\mathcal{L}{\cap}}$ is closed on projection by
    a lexicon, we have that
    $\mathcal{L}_{(u,v_n)}(\mathcal{E}_{\mathcal{G}}(u)) \in
    \mathcal{I}^{\mathcal{L}{\cap}}$ as well. Finally, by the closure of
    $\mathcal{I}^{\mathcal{L}{\cap}}$ on intersection, we have that their
    intersection $\bigcap_{(u,v_n) \in E(G)}
    \mathcal{L}_{(u,v_n)}(\mathcal{E}_{\mathcal{G}}(u))$ also belongs to
    $\mathcal{I}^{\mathcal{L}{\cap}}$. Since $\mathcal{I}_{\mathcal{G}}(v_n)$
    belongs to $\mathcal{I}^{\mathcal{L}{\cap}}$ too, we know that their
    intersection, $\mathcal{E}_{\mathcal{G}}(v_n)$ belongs to
    $\mathcal{I}^{\mathcal{L}{\cap}}$.
  \end{itemize}

  Now we prove the converse, that is that $\mathcal{I}^{\mathcal{L}{\cap}}
  \subseteq \mathcal{E}$. We do so by proving that $\mathcal{I} \subseteq
  \mathcal{E}$ and that $\mathcal{E}$ is closed on projection through a
  lexicon and intersection.

  \begin{description}
    \item[Inclusion of intrinsic languages] Let us consider the intrinsic
      language $\mathcal{I}_{\mathcal{G}}(u)$ of some node $u$ in some G-ACG
      $\mathcal{G}$. We can produce a G-ACG $\mathcal{G}_u$, whose graph
      contains only the node $u$ labeled with the same signature and
      distinguished type as in $\mathcal{G}$. Since $u$ has no predecessors in
      $\mathcal{G}_u$, it will be the case that
      $\mathcal{E}_{\mathcal{G}_u}(u) = \mathcal{I}_{\mathcal{G}_u}(u) =
      \mathcal{I}_{\mathcal{G}}(u)$ and therefore the intrinsic language
      $\mathcal{I}_{\mathcal{G}}(u)$ is also an extrinsic language
      $\mathcal{E}_{\mathcal{G}_u}(u)$.

    \vspace{5mm}

    \item[Closure on projection through a lexicon] Let $L_1$ and $L_2$ be two
      extrinsic languages defined by the nodes $u_1$ and $u_2$ in some
      graphical ACGs $\mathcal{G}_1$ and $\mathcal{G}_2$, respectively. We can
      construct a new graphical ACG $\mathcal{G}'$ by taking the union of the
      two ACGs (union of graphs (union of vertices and edges) and unions of
      the labelings).\footnote{Assuming, without loss of generality, that the
        sets of vertices of the two graphs are disjoint.} We will then add a
      new node $v$ such that $\mathcal{E}_{\mathcal{G}'}(v) = L_1 \cap L_2$.

      The signature labeling $v$ will be the union of the two signatures
      labeling $u_1$ and $u_2$, ensuring that $\forall u \in \{u_1,
      u_2\}$, $\mathcal{I}_{\mathcal{G}'}(v) \supseteq
      \mathcal{I}_{\mathcal{G}'}(u) \supseteq
      \mathcal{E}_{\mathcal{G}'}(u)$. We can use either of the two
      distinguished types that label $u_1$ or $u_2$ to label $v$ (if
      they are different, then the intersection $L_1 \cap L_2$ is
      trivially empty). Finally, we will add two edges leading from
      $u_1$ and $u_2$ to $v$, both labeled with the identity lexicon. We
      will now show that $\mathcal{E}_{\mathcal{G}'}(v) = L_1 \cap L_2$.

      \begin{align*}
        \mathcal{E}_{\mathcal{G}'}(v) &= \mathcal{I}_{\mathcal{G}'}(v)
        \cap \mathcal{L}_{(u_1,v)}(\mathcal{E}_{\mathcal{G}'}(u_1)) \cap
        \mathcal{L}_{(u_2,v)}(\mathcal{E}_{\mathcal{G}'}(u_2)) \\ &=
        \mathcal{I}_{\mathcal{G}'}(v) \cap
        \mathcal{E}_{\mathcal{G}'}(u_1) \cap
        \mathcal{E}_{\mathcal{G}'}(u_2) \\ &=
        \mathcal{E}_{\mathcal{G}'}(u_1) \cap
        \mathcal{E}_{\mathcal{G}'}(u_2) \\ &=
        \mathcal{E}_{\mathcal{G}_1}(u_1) \cap
        \mathcal{E}_{\mathcal{G}_2}(u_2) \\ &= L_1 \cap L_2
      \end{align*}

      The simplifications use the facts that: the two new lexicons are
      identities, that $\mathcal{I}_{\mathcal{G}'}(v) \supseteq
      \mathcal{E}_{\mathcal{G}'}(u_1) \cap \mathcal{E}_{\mathcal{G}'}(u_2)$
      (due to $\Sigma'_v = \Sigma'_{u_1} \cup \Sigma'_{u_2}$), that
      $\mathcal{E}_{\mathcal{G}'}(u_1) = \mathcal{E}_{\mathcal{G}_1}(u_1)$ and
      $\mathcal{E}_{\mathcal{G}'}(u_2) = \mathcal{E}_{\mathcal{G}_2}(u_2)$
      (due to $u_1$ and $u_2$ having the same predecessors in $\mathcal{G}$
      and $\mathcal{G}'$) and finally that $\mathcal{E}_{\mathcal{G}}(u_1) =
      L_1$ and $\mathcal{E}_{\mathcal{G}}(u_2) = L_2$ (our starting
      assumptions).

  \vspace{5mm}

  \item[Closure on intersection] Let $L \subseteq \Lambda(\Sigma_O)$ be a
    language which is the result of mapping an extrinsic language $L'
    \subseteq \Lambda(\Sigma_A)$ through the lexicon $\mathcal{L} : \Sigma_A
    \to \Sigma_O$. We will show that $L$ is also an extrinsic language. Let
    $u$ be the node that defines the extrinsic language $L$ in some graphical
    ACG $\mathcal{G}$. We construct a new graphical ACG $\mathcal{G}'$ by
    adding a new vertex $v$ to $\mathcal{G}$ labeled with the signature
    $\Sigma_O$ and the distinguished type $\mathcal{L}(S)$ where $S$ is the
    distinguished type labeling $u$. To this graph, we conjoin the edge
    $(u,v)$ and label it with $\mathcal{L}$.

    Now we verify that $\mathcal{E}_{\mathcal{G}'}(v) = L$.

    \begin{align*}
      \mathcal{E}_{\mathcal{G}'}(v) &= \mathcal{I}_{\mathcal{G}'}(v) \cap
      \mathcal{L}(\mathcal{E}_{\mathcal{G}'}(u)) \\
      &= \mathcal{L}(\mathcal{E}_{\mathcal{G}'}(u)) \\
      &= \mathcal{L}(\mathcal{E}_{\mathcal{G}}(u)) \\
      &= \mathcal{L}(L') \\
      &= L
    \end{align*}

    The assumptions used here are in turn: that $\mathcal{I}_{\mathcal{G}'}(v)
    \supseteq \mathcal{L}(\mathcal{E}_{\mathcal{G}'}(u))$ (due to the
    homomorphism property of lexicon $\mathcal{L} : \Sigma_A \to \Sigma_O$),
    that $\mathcal{E}_{\mathcal{G}'}(u) = \mathcal{E}_{\mathcal{G}}(u)$ (due
    to $u$ having the same predecessors in $\mathcal{G}$ and $\mathcal{G}'$),
    that $\mathcal{E}_{\mathcal{G}}(u) = L'$ and that $\mathcal{L}(L') = L$
    (our starting assumptions).
  \end{description}
\hfill $\qed$
\end{proof}

If we want our grammatical framework to be modular w.r.t. the different
linguistic constraints it encodes, we essentially need
intersection\footnote{If I have a grammar that enforces $A$ and a grammar that
  enforces $B$, then I want to have access to the grammar of the language
  where both $A$ and $B$ are enforced.}. Not only would we want a framework in
which intersection is possible, we would also like it to be an easy
operation. Graphical ACGs make intersections trivial to encode and as for
expressive power, we know, from the equations above, that object languages are
as expressive as extrinsic languages iff object languages are closed on
intersection (which is, at this moment, conjectured to be false). This means
that graphical ACGs are either a conservative extension providing a more
convenient way of expressing intersection or are extending ACGs by adding
only intersection, which enables constraint-based composition of grammars.

\begin{figure}
  \centering
  \includegraphics[height=0.3\textheight]{../diagrams/diamond-grammar.pdf}
  \caption{\label{fig:diamond} The diamond-shaped G-ACG $\mathcal{D}$.}
\end{figure}

While the interpretation of G-ACGs given above does have some nice properties,
it fails to predict the intuitive understanding of more complex ACG
diagrams. This is most visible in the diamond-shaped grammar on
Figure~\ref{fig:diamond}. If we take a term from
$\mathcal{E}_{\mathcal{D}}(Bottom)$, we know it has antecedents both in
$\mathcal{E}_{\mathcal{D}}(Left)$ and
$\mathcal{E}_{\mathcal{D}}(Right)$. However, these two do not have to share an
antecedent in $\mathcal{E}_{\mathcal{D}}(Top)$. This contradicts the
generative story one might imagine in which a single term from $Top$ generates
terms in $Left$ and $Right$ which generate a term in $Bottom$.

We can observe another peculiarity on a more practical example. Consider the
G-ACG in Figure~\ref{fig:gacg}. In classical ACGs, one can always take an
element in a signature, like $Sem$, and by transduction find its corresponding
element in an another signature, like $Syntax$. However, in a G-ACG such as
this, it is possible that the $Syntax$ term we obtain by transduction does not
belong to $\mathcal{E}_{\mathcal{G}}(Syntax)$ because it lacks antecedents in
either $Constr_1$ or $Constr_2$. This means that the notion of an extrinsic
language is not capable to give us the set of all meaning terms in $Sem$ which
actually correspond to syntactically correct sentences in this G-ACG.

The above characteristics motivated us to explore alternative interpretations
of G-ACGs. We will present one such alternative now, which exchanges the
language-algebraic point of view for a term-relational one.

\subsection{Nodes as Terms}

In the new paradigm, we interpret the graph as a logical formula. The nodes
will be seen as variables whose domains are their intrinsic languages,
i.e. each node $v$ as a variable will range over the terms built from its
signature $\Sigma_v$ and having its distinguished type $S_v$. Edges $(v,w)$
will be statements saying that the image of the term $v$ through the lexicon
$\mathcal{L}_{(v,w)}$ is exactly the term $w$.

We can think of \emph{models} for this graphical propositional logic as
assignments of terms to variables, i.e. labelings of the nodes of the graph
with terms. A model $T$ is \emph{valid} when it satisfies the following
conditions:

\begin{itemize}
  \item For all $v \in V(G)$, $\vdash_{\Sigma_v} T_v : S_v$.
  \item For all $(v,w) \in E(G)$, $\mathcal{L}_{(v,w)}(T_v) = T_w$.
\end{itemize}

A term $t$ belongs to the \emph{pangraphical}\footnote{As opposed to extrinsic
  languages, which are constrained only by their predecessors in the graph,
  pangraphical languages are constrained by the entire graph, hence the name.}
language $\mathcal{P}_{\mathcal{G}}(u)$ if there exists a valid model $T$ such
that $T(u) = t$.

If we compare this new interpretation of G-ACGs to the former one, we find out
that in the cases when the graph of the grammar is an arborescence, they are
actually equivalent. This means that in classical ACGs, where all the diagrams
are arborescences, the two metaphors (nodes as languages and nodes as terms)
can be, and are, used interchangeably. However, as we start to consider
non-arborescent graphs, we find, interestingly, that the two paradigms diverge
(i.e. $\exists \mathcal{G}, u.\ \mathcal{E}_{\mathcal{G}}(u) \neq
\mathcal{P}_{\mathcal{G}}(u)$).

The newly defined pangraphical languages solve the problem of extrinsic
languages giving us counter-intuitive interpretations for some specific
G-ACGs. Specifically, the members of $\mathcal{P}_{\mathcal{D}}(Bottom)$ in
the diamond grammar have a single antecedent in
$\mathcal{P}_{\mathcal{D}}(Top)$ and the language
$\mathcal{P}_{\mathcal{G}}(Sem)$ (from Figure~\ref{fig:gacg}) contains only
meanings expressible by syntactically correct sentences.

Pangraphical languages turn out to be at least as expressive as the extrinsic
languages\footnote{This might come as no surprise given the extended domain of
  constraints compared to extrinsic languages (constrained by the entire graph
  as opposed to just the node's predecessors).}. The proof is carried out by
transforming any G-ACG $\mathcal{G}$ where $L = \mathcal{E}_{\mathcal{G}}(u)$
to a G-ACG $\mathcal{G'}$ where for some $v$, $\mathcal{P}_{\mathcal{G'}}(v) =
L$. We do so by taking all the (finitely many) paths in $G$ that end in $u$ to
be the vertices of $G'$. Each of the vertices in $G'$ (paths in $G$) will be
labeled by the same signature and distinguished type as the first node of the
corresponding path. Furthermore, there will be an edge $(p_1, p_2)$ in $G'$
labeled with the lexicon $\mathcal{L}$ whenever the path $p_2$ is the maximal
strict suffix of $p_1$ such that $p_1 = e; p_2$ where the edge $e$ is labeled
with $\mathcal{L}$ in $G$.\footnote{This means that when the G-ACG is
  $\mathcal{D}$, the diamond grammar, and the node in question is $Bottom$,
  then we will have the following nodes in $\mathcal{D}'$: $[Top, Left,
    Bottom]$, $[Top, Right, Bottom]$, $[Left, Bottom]$, $[Right, Bottom]$,
  $[Bottom]$. Note that we have two nodes for (i.e. beginning with) $Top$,
  which will let us admit two different antecedents in
  $\mathcal{E}_{\mathcal{D}}(Top)$.} In this G-ACG $\mathcal{G'}$, valid
models $T$ which assign $T([u]) = t$ are isomorphic to proofs of membership of
$t$ in $\mathcal{E}_{\mathcal{G}}(u)$ and therefore
$\mathcal{P}_{\mathcal{G'}}([u]) = \mathcal{E}_{\mathcal{G}}(u)$.

This gives us the following ladder of expressivity

$$
\mathcal{I} \subseteq \mathcal{E} \subseteq \mathcal{P}
$$

which can be complemented with the following series

$$
\mathcal{I}_{\mathcal{G}}(u) \supseteq \mathcal{E}_{\mathcal{G}}(u) \supseteq
\mathcal{P}_{\mathcal{G}}(u)
$$

\section{Illustration}

In this final section, we assemble a G-ACG which integrates the French
negation constraint discussed in Section~\ref{sec:constraints}, the
constraints on extraction introduced in \cite{pogodalla2012controlling} and a
constraint handling agreement in a single grammar specification to demonstrate
our approach.

\begin{figure}[t]
  \centering
  \begin{subfigure}[b]{0.4\textwidth}
    \centering
    \includegraphics[height=0.2\textheight]{../diagrams/story-single-cont.pdf}
    \caption{\label{fig:single-cont} An extraction constraint }
  \end{subfigure}
  \qquad
  \begin{subfigure}[b]{0.4\textwidth}
    \centering
    \includegraphics[height=0.2\textheight]{../diagrams/story-merged-conts.pdf}
    \caption{\label{fig:merged-conts} G-ACG merging the three extraction
      constraints.}
  \end{subfigure}
  \caption{\label{fig:conts} Combining the constraints on extraction.}
\end{figure}

We start with the extraction constraints. \cite{pogodalla2012controlling}
describes three different constraints on extraction, all of them expressed
using the Constraint pattern (Figure~\ref{fig:single-cont}). We can take the
union of these three G-ACGs to get the G-ACG on Figure~\ref{fig:merged-conts}.
Next, we can incorporate knowledge from \cite{pogodalla2007generalizing}, by
splitting the $\mathcal{L}_{Syn}$ lexicon and introducing a new intermediate
signature (Figure~\ref{fig:conts-elab}). The new signature, $Syntax$, deals
purely with syntax without any issues of operator scope and its functions have
types of a lower order than those in $Syn$. We will finally add our
contribution, the signatures $Neg$ and $Agr$ and the lexicons
$\mathcal{L}_{Neg}$ and $\mathcal{L}_{Agr}$ which implement the constraint on
negation we discussed before and some simple notion of agreement, respectively
(Figure~\ref{fig:final-synth}).

\begin{figure}[t]
  \centering
  \begin{subfigure}[b]{0.4\textwidth}
    \centering
    \includegraphics[height=0.2\textheight]{../diagrams/story-conts-elab.pdf}
    \caption{\label{fig:conts-elab} Elaborating the lexicon from syntax to
      strings. }
  \end{subfigure}
  \qquad
  \begin{subfigure}[b]{0.4\textwidth}
    \centering
    \includegraphics[height=0.2\textheight]{../diagrams/story-final-synth.pdf}
    \caption{\label{fig:final-synth} Adding our own syntactic constraints,
      $Neg$ and $Agr$.}
  \end{subfigure}
  \caption{\label{fig:final} Building the example grammar up to its final
    state.}
\end{figure}

We notice that while the three constraints introduced in
\cite{pogodalla2012controlling} are expressed in terms of the $Syn$ signature,
our constraints $Agr$ and $Neg$ use a different interface, the $Syntax$
signature, to constrain the languages being defined. The $Cont_i$ constraints
are not actually constraining syntax itself, but the syntax-semantics
interface. This is in some cases a deliberate decision, since e.g. $Cont_1$ is
the constraint enforcing that quantifier scope must not escape outside of
tensed clauses. The $Syntax$ signature is not equipped to talk about operator
scope and so the constraint is expressed above the $Syn$ signature. On the
other hand, constraints $Cont_2$ and $Cont_3$ from
\cite{pogodalla2012controlling} are purely syntactic and therefore could be
more easily expressed in terms of $Syntax$ instead of $Syn$. However, we see
it as a perk of our approach that we can ignore that and import the
constraints wholesale without having to adapt them.

\subsection{The $Neg$ Signature}

We will now look under the covers of the two signatures implementing the
constraints, starting with negation and the $Neg$ signature.

As was said in Section~\ref{sec:constraints}, we will proceed by splitting the
atomic types $N$ and $NP$ into $N_{NEG{=}F}$ and $N_{NEG{=}T}$, and
$NP_{NEG{=}F}$ and $NP_{NEG{=}T}$, respectively. We will use these new types
to distinguish whether an $N$ or $NP$ phrase contains a negative determiner
that can pair up with a negated verb. Where before we had a single type, we
will now have possibly multiple types to account for the different values of
the arguments' $NEG$ features. Here are some representative examples:

\begin{align*}
N_{aucun} &: N_{NEG{=}F} \limp NP_{NEG{=}T} \\
\end{align*} % for page break purposes
\begin{align*}
N_{le_1} &: N_{NEG{=}F} \limp NP_{NEG{=}F} \\
N_{le_2} &: N_{NEG{=}T} \limp NP_{NEG{=}T} \\
N_{ne^{tv}_1} &: (NP_{NEG{=}F} \limp NP_{NEG{=}F} \limp S) \limp (NP_{NEG{=}T} \limp NP_{NEG{=}F} \limp S) \\
N_{ne^{tv}_2} &: (NP_{NEG{=}F} \limp NP_{NEG{=}F} \limp S) \limp (NP_{NEG{=}F}
\limp NP_{NEG{=}T} \limp S) \\
N_{ne^{tv}_3} &: (NP_{NEG{=}F} \limp NP_{NEG{=}F} \limp S) \limp (NP_{NEG{=}T}
\limp NP_{NEG{=}T} \limp S) \\
N_{aime} &: NP_{NEG{=}F} \limp NP_{NEG{=}F} \limp S \\
N_{que_1} &: (NP_{NEG{=}F} \limp S) \limp N_{NEG{=}F} \limp N_{NEG{=}F} \\
N_{que_2} &: (NP_{NEG{=}F} \limp S) \limp N_{NEG{=}T} \limp N_{NEG{=}T} \\
\end{align*}

\vspace{-3ex}

We see that the negative determiner \emph{aucun} can only attach to phrases
which do not already contain a free negative determiner and that the resulting
phrase is marked as having a free negative determiner. On the other hand, the
determiner \emph{le} does not have any interaction with the new $NEG$ feature
and is still able to take any $N$ phrase, preserving any free negative
determiners inside. The type of the negative particle \emph{ne} modifies a
verb (in the example above, $N_{ne^{tv}_i}$, a transitive verb) and produces a
new verb that makes sure that at least one of its arguments contains a
negative determiner. Verbs are not agnostic about the $NEG$ feature. By
default, they specifically demand it to be false, since a negative determiner
cannot be coupled with a non-negated verb. Finally, the types of the
relativizer \emph{que} tell us two other things: the $NP$ trace is considered
as containing no negative determiners, and a relative clause does not care
about or alter the presence of free negative determiners in the $N$ it
modifies.

The lexicon which translates from this signature to $Syntax$ is a
straightforward relabeling. Its items can be characterized by the following
schema:

$$
\mathcal{L}_{Neg}(N_{wordform_{i}}) = C_{wordform}
$$

where $C_{wordform}$ is the constant in $Syntax$ corresponding to the
$wordform$.

\subsection{The $Agr$ Signature}

The $Agr$ signature will be constructed using the same strategy as $Neg$. $N$
and $NP$ will be subdivided into $N_{NUM{=}SG}$, $N_{NUM{=}PL}$,
$NP_{NUM{=}SG}$ and $NP_{NUM{=}PL}$ (we only treat number agreement in this
example). Here are some example types:

\begin{align*}
A_{tatou} &: N_{NUM{=}SG} \\
A_{tatous} &: N_{NUM{=}PL} \\
A_{le} &: N_{NUM{=}SG} \limp NP_{NUM{=}SG} \\
A_{les} &: N_{NUM{=}PL} \limp NP_{NUM{=}PL} \\
A_{aime_1} &: NP_{NUM{=}SG} \limp NP_{NUM{=}SG} \limp S \\
A_{aime_2} &: NP_{NUM{=}SG} \limp NP_{NUM{=}PL} \limp S \\
\end{align*} % for page break purposes
\begin{align*}
A_{qui_1} &: (NP_{NUM{=}SG} \limp S) \limp N_{NUM{=}SG} \limp N_{NUM{=}SG} \\
A_{qui_2} &: (NP_{NUM{=}PL} \limp S) \limp N_{NUM{=}PL} \limp N_{NUM{=}PL} \\
\end{align*}

The types for nouns and determiners are quite predictable. For the transitive
verb \emph{aime}, we see that we need a second type to account for the fact
that \emph{aime} does not care about the number of the object. Finally, the
relativizer \emph{qui} enforces that the number of the trace $NP$ in the
relative clause must be the same as the number of the $N$ being modified.

The lexicon $\mathcal{L}_{Agr}$ follows exactly the same schema as that of
$\mathcal{L}_{Neg}$.

\subsection{Limits of Modularity}

We have decomposed our grammar into 9 signatures and 8 lexicons. Besides
making the grammar more readable and easily navigable, it also makes the
grammar more attractive from a computational point of view. The complexity of
parsing the highest abstract-most structures and verifying all the constraints
is decomposed into smaller subtasks. A string can be parsed into its
representation at the $Syntax$ level fairly easily (compared to parsing it at
the higher levels) since $Syntax$ employs types of lower orders than the more
abstract signatures. Finding the syntactic structure for a string is thus
feasible. We then proceed to verifying the syntactic constraints. Here we can
observe that the lexicons $\mathcal{L}_{Neg}$ and $\mathcal{L}_{Agr}$ are mere
relabelings and therefore parsing (i.e. constraint verification) is trivial
(i.e. decidable). Now in order to determine the meaning of our sentence, we
have to continue parsing by finding an antecedent in $Syn$. This involves
higher-order types used to encode the various scope-bearing
operators. Finally, we can opt in to verify the high-level syntax-semantics
constraints which employ dependent types
\cite{pogodalla2012controlling}. However, the lexicons in this case resemble
relabelings and a semi-decidable procedure could be used to solve the
problem. Having the grammar decomposed like this opens it up for computational
processing by letting us engage in different tactics at different stages
instead of having to parse a single hidden structure which enforces
grammaticality w.r.t. every constraint. G-ACGs share these perks also with
classical ACGs but they let use this composition in another dimension, which
facilitates the writing of multiple constraints.

However, it is important to keep in mind that the composability demonstrated
above is limited only to the composition of constraints. If we would like to
extend our language with, e.g., a new lexical category, then every signature
and every lexicon would have to be extended to cover this new category. It is
also important to point out that the constraint signatures that we have been
composing are not light-weight objects. They are usually defined by taking an
existing signature that describes the language at a relevant level and by
strengthening (usually all) the types of that signature to enforce the
constraint\footnote{This makes the expressivity of model-theoretic syntactic
  formalisms such as Interaction Grammars, in which adding the lexical items
  for \emph{ne} and \emph{aucun} along with enforcing the $Neg$ constraint
  takes only two tree descriptions \cite{perrier2007french}, even more
  striking when compared to our approach.}. However, most of these type
alterations are mundane and we believe that this is fertile ground for
applying metagrammars.

\section{Conclusion}

We have considered the problem of building a wide-coverage ACG, specifically
the problem of expressing a multitude of linguistic constraints. We have
examined previous techniques and found no satisfying solution. We have thus
provided an extension of the ACG formalism to solve the problem and justified
the need for the increased expressivity. This embedding of syntactic
constraints will contribute to an effort to define a syntax-semantics
interface and later to build discourse structures.

In the end, our approach lets us define the syntax in a clean way using the
idiomatic style of categorial grammars (simple atomic types like $N$, $NP$ and
$S$) and then define the constraints themselves the same way as they are
defined in ACG research (such as is the case with
\cite{pogodalla2012controlling}).

Interesting avenues of future work include: digging deeper into the
language-theoretical properties of G-ACGs (Are ACG object languages closed on
intersection? Are extrinsic languages as expressive as pangraphical
languages?), searching for a metagrammatical description of constraint
signatures to ease work on G-ACG grammars, and finally building large-scale
grammars to verify the usability of our approach.

%
% ---- Bibliography ----
%

\bibliographystyle{splncs}
\bibliography{../biblio}

\end{document}